---
title: "Exp2_Analysis"
author: "Will Hayes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Experiment 2 Analyses

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(afex)
```

## Load data

1. Load the empirical data

```{r}
source('../modeling/load_data.R')
```


2. Load the model simulations

```{r}
sims1 <- readRDS('../modeling/results/model1_sims.RDS')
sims2 <- readRDS('../modeling/results/model2_sims.RDS')
sims3 <- readRDS('../modeling/results/model3_sims.RDS')
sims4 <- readRDS('../modeling/results/free-theta/model4_sims.RDS')
sims5 <- readRDS('../modeling/results/free-theta/model5_sims.RDS')
sims6 <- readRDS('../modeling/results/free-theta/model6_sims.RDS')
sims7 <- readRDS('../modeling/results/free-theta/model7_sims.RDS')
sims8 <- readRDS('../modeling/results/free-theta/model8_sims.RDS')
```


Model Version 2 includes a trial-dependent (decreasing) decision threshold

```{r}
sims1_v2 <- readRDS('../modeling/results/tdep_thresh_model1_sims.RDS')
sims2_v2 <- readRDS('../modeling/results/tdep_thresh_model2_sims.RDS')
sims3_v2 <- readRDS('../modeling/results/tdep_thresh_model3_sims.RDS')
sims4_v2 <- readRDS('../modeling/results/free-theta/tdep_thresh_model4_sims.RDS')
sims5_v2 <- readRDS('../modeling/results/free-theta/tdep_thresh_model5_sims.RDS')
sims6_v2 <- readRDS('../modeling/results/free-theta/tdep_thresh_model6_sims.RDS')
sims7_v2 <- readRDS('../modeling/results/free-theta/tdep_thresh_model7_sims.RDS')
sims8_v2 <- readRDS('../modeling/results/free-theta/tdep_thresh_model8_sims.RDS')
```


## Set order of models for plotting

```{r}
model_order <- c('model_1', 'model_2', 'model_3', 'model_4',
                 'model_5', 'model_6', 'model_7', 'model_8')

model_labels <- c('Q',
                  'Q * gaze',
                  'Q + gaze',
                  'softmax(Q)',
                  'softmax(Q * gaze)',
                  'softmax(Q) * gaze',
                  'softmax(Q) + gaze',
                  'softmax(Q + gaze)')
```


## Learning curves

1. Empirical learning curve

```{r}
# compute proportion of correct choices on each trial (across participants)
lcData <- dat %>%
  filter(block == 'learning') %>%
  group_by(trial_number) %>%
  summarise(acc = mean(correct),
            n = n()) %>%
  ungroup() %>%
  mutate(se = sqrt(acc * (1 - acc) / n),
         upper = acc + se,
         lower = acc - se)
```


2. Model-simulated learning curves

```{r}
# Step 1: For each subject, compute p(correct) on each learning phase trial (1:120) across the 100 simulation runs
# Returns a 120 (trials) x 50 (subjects) matrix
model1_lc <- sapply(sims1, function(X) rowMeans(X[1:120,1,]), simplify='array')
model2_lc <- sapply(sims2, function(X) rowMeans(X[1:120,1,]), simplify='array')
model3_lc <- sapply(sims3, function(X) rowMeans(X[1:120,1,]), simplify='array')
model4_lc <- sapply(sims4, function(X) rowMeans(X[1:120,1,]), simplify='array')
model5_lc <- sapply(sims5, function(X) rowMeans(X[1:120,1,]), simplify='array')
model6_lc <- sapply(sims6, function(X) rowMeans(X[1:120,1,]), simplify='array')
model7_lc <- sapply(sims7, function(X) rowMeans(X[1:120,1,]), simplify='array')
model8_lc <- sapply(sims8, function(X) rowMeans(X[1:120,1,]), simplify='array')

model1_v2_lc <- sapply(sims1_v2, function(X) rowMeans(X[1:120,1,]), simplify='array')
model2_v2_lc <- sapply(sims2_v2, function(X) rowMeans(X[1:120,1,]), simplify='array')
model3_v2_lc <- sapply(sims3_v2, function(X) rowMeans(X[1:120,1,]), simplify='array')
model4_v2_lc <- sapply(sims4_v2, function(X) rowMeans(X[1:120,1,]), simplify='array')
model5_v2_lc <- sapply(sims5_v2, function(X) rowMeans(X[1:120,1,]), simplify='array')
model6_v2_lc <- sapply(sims6_v2, function(X) rowMeans(X[1:120,1,]), simplify='array')
model7_v2_lc <- sapply(sims7_v2, function(X) rowMeans(X[1:120,1,]), simplify='array')
model8_v2_lc <- sapply(sims8_v2, function(X) rowMeans(X[1:120,1,]), simplify='array')

# Step 2: Average across subjects
# Returns a vector of simulated mean p(correct) on each trial
model1_lc <- rowMeans(model1_lc)
model2_lc <- rowMeans(model2_lc)
model3_lc <- rowMeans(model3_lc)
model4_lc <- rowMeans(model4_lc)
model5_lc <- rowMeans(model5_lc)
model6_lc <- rowMeans(model6_lc)
model7_lc <- rowMeans(model7_lc)
model8_lc <- rowMeans(model8_lc)

model1_v2_lc <- rowMeans(model1_v2_lc)
model2_v2_lc <- rowMeans(model2_v2_lc)
model3_v2_lc <- rowMeans(model3_v2_lc)
model4_v2_lc <- rowMeans(model4_v2_lc)
model5_v2_lc <- rowMeans(model5_v2_lc)
model6_v2_lc <- rowMeans(model6_v2_lc)
model7_v2_lc <- rowMeans(model7_v2_lc)
model8_v2_lc <- rowMeans(model8_v2_lc)
```


3. Merge model predictions with the empirical data.

```{r}
lcData2 <- lcData

lcData <- lcData %>%
  mutate(model_1 = model1_lc,
         model_2 = model2_lc,
         model_3 = model3_lc,
         model_4 = model4_lc,
         model_5 = model5_lc,
         model_6 = model6_lc,
         model_7 = model7_lc,
         model_8 = model8_lc)

lcData2 <- lcData2 %>%
  mutate(model_1 = model1_v2_lc,
         model_2 = model2_v2_lc,
         model_3 = model3_v2_lc,
         model_4 = model4_v2_lc,
         model_5 = model5_v2_lc,
         model_6 = model6_v2_lc,
         model_7 = model7_v2_lc,
         model_8 = model8_v2_lc)

# reshape to long format
lcData1 <- lcData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

lcData3 <- lcData2 %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# turn model into a factor to control ordering
lcData1 <- lcData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))

# add model version 2 predictions 
lcData1$fit_v2 <- lcData3$fit
```


WRITE DATA TO FILE

```{r}
#write.csv(lcData1, 'Exp2_LearningCurve_Fits.csv', row.names=F)
```


4. Plot of learning curve fits for all models

```{r}
p1 <- ggplot(lcData1, aes(x=trial_number, y=acc)) +
  facet_wrap( ~model, nrow=1) +
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3) +
  geom_line(aes(y=fit), col='maroon1', lty=1) +
  geom_line(aes(y=fit_v2), col='orange1', lty=1) +
  ylab('Choice accuracy') +
  xlab('Trial number') +
  ggpubr::theme_pubclean() +
  ggtitle('Exp. 2 Learning Phase: Accuracy Curve') +
  theme(plot.title = element_text(hjust=0.5)) +
  scale_x_continuous(breaks=seq(30, 120, 30))
```


PLOT FOR MAIN MANUSCRIPT (Fig 4E)

```{r}
# png('Exp2_LearningCurve.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(lcData1, model %in% c('Q + gaze', 'softmax(Q + gaze)')),
#        aes(x=trial_number, y=acc)) +
#   geom_line() +
#   geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3) +
#   geom_line(aes(y=fit, col=model), lty=1, lwd=0.8) +
#   ylab('Choice accuracy') +
#   xlab('Trial number') +
#   scale_x_continuous(breaks=seq(0, 120, 30)) +
#   scale_color_manual(values=c('green', 'maroon1')) +
#   theme_classic(base_size = 11) +
#   theme(panel.grid.minor.x = element_blank(),
#         panel.grid.major.x = element_blank(),
#         panel.grid.minor.y = element_blank(),
#         panel.grid.major.y = element_blank(),
#         legend.position = 'inside',
#         legend.position.inside = c(.65, .2),
#         legend.title = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#     ggtitle('Exp. 2: Learning')
# dev.off()
```



## Response time curves


1. Empirical response time curve (RTs < 250 ms or > 10 sec excluded)

```{r}
# compute mean RT on each trial (across participants)
rtcData <- dat %>%
  filter(block == 'learning', RT >= 250, RT <= 10000) %>%
  group_by(trial_number) %>%
  summarise(mean_RT = mean(RT),
            sd_RT = sd(RT),
            n = n()) %>%
  ungroup() %>%
  mutate(se = sd_RT / sqrt(n),
         upper = mean_RT + se,
         lower = mean_RT - se)
```


2. Model-simulated RT curves

```{r}
# Step 1: For each subject, compute mean RT on each learning phase trial (1:120) across the 100 simulation runs
# Returns a 120 (trials) x 50 (subjects) matrix
model1_rtc <- sapply(sims1, function(X) rowMeans(X[1:120,2,]), simplify='array')
model2_rtc <- sapply(sims2, function(X) rowMeans(X[1:120,2,]), simplify='array')
model3_rtc <- sapply(sims3, function(X) rowMeans(X[1:120,2,]), simplify='array')
model4_rtc <- sapply(sims4, function(X) rowMeans(X[1:120,2,]), simplify='array')
model5_rtc <- sapply(sims5, function(X) rowMeans(X[1:120,2,]), simplify='array')
model6_rtc <- sapply(sims6, function(X) rowMeans(X[1:120,2,]), simplify='array')
model7_rtc <- sapply(sims7, function(X) rowMeans(X[1:120,2,]), simplify='array')
model8_rtc <- sapply(sims8, function(X) rowMeans(X[1:120,2,]), simplify='array')

model1_v2_rtc <- sapply(sims1_v2, function(X) rowMeans(X[1:120,2,]), simplify='array')
model2_v2_rtc <- sapply(sims2_v2, function(X) rowMeans(X[1:120,2,]), simplify='array')
model3_v2_rtc <- sapply(sims3_v2, function(X) rowMeans(X[1:120,2,]), simplify='array')
model4_v2_rtc <- sapply(sims4_v2, function(X) rowMeans(X[1:120,2,]), simplify='array')
model5_v2_rtc <- sapply(sims5_v2, function(X) rowMeans(X[1:120,2,]), simplify='array')
model6_v2_rtc <- sapply(sims6_v2, function(X) rowMeans(X[1:120,2,]), simplify='array')
model7_v2_rtc <- sapply(sims7_v2, function(X) rowMeans(X[1:120,2,]), simplify='array')
model8_v2_rtc <- sapply(sims8_v2, function(X) rowMeans(X[1:120,2,]), simplify='array')

# Step 2: Average across subjects
# Returns a vector of simulated mean RTs on each trial
model1_rtc <- rowMeans(model1_rtc)
model2_rtc <- rowMeans(model2_rtc)
model3_rtc <- rowMeans(model3_rtc)
model4_rtc <- rowMeans(model4_rtc)
model5_rtc <- rowMeans(model5_rtc)
model6_rtc <- rowMeans(model6_rtc)
model7_rtc <- rowMeans(model7_rtc)
model8_rtc <- rowMeans(model8_rtc)

model1_v2_rtc <- rowMeans(model1_v2_rtc)
model2_v2_rtc <- rowMeans(model2_v2_rtc)
model3_v2_rtc <- rowMeans(model3_v2_rtc)
model4_v2_rtc <- rowMeans(model4_v2_rtc)
model5_v2_rtc <- rowMeans(model5_v2_rtc)
model6_v2_rtc <- rowMeans(model6_v2_rtc)
model7_v2_rtc <- rowMeans(model7_v2_rtc)
model8_v2_rtc <- rowMeans(model8_v2_rtc)
```


3. Merge model predictions with the empirical data.

```{r}
rtcData2 <- rtcData

rtcData <- rtcData %>%
  mutate(model_1 = model1_rtc,
         model_2 = model2_rtc,
         model_3 = model3_rtc,
         model_4 = model4_rtc,
         model_5 = model5_rtc,
         model_6 = model6_rtc,
         model_7 = model7_rtc,
         model_8 = model8_rtc)

rtcData2 <- rtcData2 %>%
  mutate(model_1 = model1_v2_rtc,
         model_2 = model2_v2_rtc,
         model_3 = model3_v2_rtc,
         model_4 = model4_v2_rtc,
         model_5 = model5_v2_rtc,
         model_6 = model6_v2_rtc,
         model_7 = model7_v2_rtc,
         model_8 = model8_v2_rtc)

# reshape to long format
rtcData1 <- rtcData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

rtcData3 <- rtcData2 %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# turn model into a factor to control ordering
rtcData1 <- rtcData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))

# add model version 2 predictions 
rtcData1$fit_v2 <- rtcData3$fit
```


WRITE DATA TO FILE

```{r}
#write.csv(rtcData1, 'Exp2_RTCurve_Fits.csv', row.names=F)
```


4. Plot of RT curve fits for all models

```{r}
p2 <- ggplot(rtcData1, aes(x=trial_number, y=mean_RT)) +
  facet_wrap( ~model, nrow=1) +
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3) +
  geom_line(aes(y=fit), col='maroon1', lty=1) +
  geom_line(aes(y=fit_v2), col='orange1', lty=1) + 
  ylab('Mean RT (ms)') +
  xlab('Trial number') +
  ggpubr::theme_pubclean()  +
  ggtitle('Exp. 2 Learning Phase: Mean Response Time Curve') +
  theme(plot.title = element_text(hjust=0.5)) +
  scale_x_continuous(breaks=seq(30, 120, 30))
```

PLOT FOR MAIN MANUSCRIPT (Fig 4F)

```{r}
# png('Exp2_RTCurve.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(rtcData1, model %in% c('Q + gaze', 'softmax(Q + gaze)')),
#        aes(x=trial_number, y=mean_RT)) +
#   geom_line() +
#   geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3) +
#   geom_line(aes(y=fit, col=model), lty=1, lwd=0.8, show.legend=FALSE) +
#   ylab('Mean RT (ms)') +
#   xlab('Trial number') +
#   scale_x_continuous(breaks=seq(0, 120, 30)) +
#   scale_color_manual(values=c('green', 'maroon1')) +
#   theme_classic(base_size = 11) +
#   theme(panel.grid.minor.x = element_blank(),
#         panel.grid.major.x = element_blank(),
#         panel.grid.minor.y = element_blank(),
#         panel.grid.major.y = element_blank(),
#         legend.position = 'inside',
#         legend.position.inside = c(.65, .2),
#         legend.title = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   ggtitle('Exp. 2: Learning')
# dev.off()
```


PLOT FOR SI APPENDIX (S7 Fig)

```{r}
# png('Exp2_ChoiceRT_All_Models.png', width=11.5, height=5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# grid.arrange(p1, p2, nrow=2, ncol=1)
# dev.off()
```


## Individual choice accuracies: Learning phase

1. Empirical choice accuracies (learning phase)

```{r}
# accuracy for each participant
accData <- dat %>%
  filter(block == 'learning') %>%
  group_by(subject) %>%
  summarise(acc = mean(correct)) %>% 
  ungroup()
```


2. Model-simulated accuracies (learning phase)

```{r}
model1_acc <- sapply(sims1, function(X) mean(X[1:120,1,]))
model2_acc <- sapply(sims2, function(X) mean(X[1:120,1,]))
model3_acc <- sapply(sims3, function(X) mean(X[1:120,1,]))
model4_acc <- sapply(sims4, function(X) mean(X[1:120,1,]))
model5_acc <- sapply(sims5, function(X) mean(X[1:120,1,]))
model6_acc <- sapply(sims6, function(X) mean(X[1:120,1,]))
model7_acc <- sapply(sims7, function(X) mean(X[1:120,1,]))
model8_acc <- sapply(sims8, function(X) mean(X[1:120,1,]))

modelAccData <- data.frame(subject = IDs,
                           model_1 = model1_acc,
                           model_2 = model2_acc,
                           model_3 = model3_acc,
                           model_4 = model4_acc,
                           model_5 = model5_acc,
                           model_6 = model6_acc,
                           model_7 = model7_acc,
                           model_8 = model8_acc)
```


3. Merge model predictions with empirical data

```{r}
accData <- accData %>%
  left_join(modelAccData, by='subject')

# reshape to long format
accData1 <- accData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# turn model into a factor to control ordering
accData1 <- accData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))
```


WRITE DATA TO FILE

```{r}
#write.csv(accData1, 'Exp2_Accuracy_Fits.csv', row.names=F)
```


4. Plot of empirical vs. fitted choice accuracies

```{r}
p1 <- ggplot(accData1, aes(x=acc, y=fit)) +
  facet_wrap( ~model, nrow=1) +
  geom_abline(slope = 1, intercept=0, lty=3) +
  geom_smooth(method='lm', col='maroon1') +
  geom_point(pch=21, fill='grey', alpha=0.5) +
  ylab('Choice accuracy (model)') +
  xlab('Choice accuracy (data)') +
  ggtitle('Learning phase') +
  ggpubr::theme_pubclean() +
  theme(panel.grid.major.y = element_blank()) +
  ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3)  +
  ggtitle('Exp. 2 Learning Phase: Individual Choice Accuracies') +
  theme(plot.title = element_text(hjust=0.5))

```

PLOT FOR MAIN MANUSCRIPT (Fig 4G)

```{r}
# png('Exp2_Accuracy.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(accData1, model=='softmax(Q + gaze)'), aes(x=acc, y=fit)) +
#   geom_abline(slope = 1, intercept=0, lty=3) +
#   geom_smooth(method='lm', col='maroon1') +
#   geom_point(pch=21, fill='grey', alpha=0.5) +
#   ylab('Choice accuracy (model)') +
#   xlab('Choice accuracy (data)') +
#   theme_classic(base_size = 11) +
#   theme(panel.grid.major.y = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3) +
#   ggtitle('Exp. 2: Learning')
# dev.off()
```


5. Relationships with model parameters (S2 Table)

```{r}
# load model 8 fits
model8_fits <- readRDS('../modeling/results/free-theta/model8_fits.RDS')

# extract parameter estimates
params <- data.frame(t(sapply(model8_fits, function(X) X$optim$bestmem)))
names(params) <- c('t0','learn_rate','Q_beta','gaze_beta','threshold_sep','upper_bound','w_rel','theta')
params$threshold <- params$upper_bound + params$threshold_sep
params$theta_times_gaze_beta <- params$theta * params$gaze_beta

# add subject ID column
params$subject <- IDs

# merge with data
accData <- accData %>%
  left_join(params, by='subject')

# multiple regression predicting choice accuracy from model parameters
summary(lm(acc ~ w_rel + learn_rate + Q_beta + gaze_beta + theta + upper_bound + threshold + t0, data=accData))
```



## Individual mean RTs: Learning phase

1. Empirical mean RTs (learning phase)

```{r}
# mean RTs for each participant
rtData <- dat %>%
  filter(block == 'learning', RT >= 250, RT <= 10000) %>%
  group_by(subject) %>%
  summarise(mean_RT = mean(RT)) %>% 
  ungroup()
```


2. Model-simulated mean RTs

```{r}
model1_rt <- sapply(sims1, function(X) mean(X[1:120,2,]))
model2_rt <- sapply(sims2, function(X) mean(X[1:120,2,]))
model3_rt <- sapply(sims3, function(X) mean(X[1:120,2,]))
model4_rt <- sapply(sims4, function(X) mean(X[1:120,2,]))
model5_rt <- sapply(sims5, function(X) mean(X[1:120,2,]))
model6_rt <- sapply(sims6, function(X) mean(X[1:120,2,]))
model7_rt <- sapply(sims7, function(X) mean(X[1:120,2,]))
model8_rt <- sapply(sims8, function(X) mean(X[1:120,2,]))

modelRTData <- data.frame(subject = IDs,
                          model_1 = model1_rt,
                          model_2 = model2_rt,
                          model_3 = model3_rt,
                          model_4 = model4_rt,
                          model_5 = model5_rt,
                          model_6 = model6_rt,
                          model_7 = model7_rt,
                          model_8 = model8_rt)
```


3. Merge model predictions with empirical data

```{r}
rtData <- rtData %>%
  left_join(modelRTData, by='subject')

# reshape to long format
rtData1 <- rtData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# turn model into a factor to control ordering
rtData1 <- rtData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))
```


WRITE DATA TO FILE

```{r}
#write.csv(rtData1, 'Exp2_MeanRT_Fits.csv', row.names=F)
```


4. Plot of empirical vs. fitted mean RTs

```{r}
p2 <- ggplot(rtData1, aes(x=mean_RT, y=fit)) +
  facet_wrap( ~model, nrow=1) +
  geom_abline(slope = 1, intercept=0, lty=3) +
  geom_smooth(method='lm', col='maroon1') +
  geom_point(pch=21, fill='grey', alpha=0.5) +
  ylab('Mean RT (model)') +
  xlab('Mean RT (data)') +
  ggtitle('Learning phase') +
  ggpubr::theme_pubclean() +
  theme(panel.grid.major.y = element_blank()) +
  ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3) +
  ggtitle('Exp. 2 Learning Phase: Individual Mean RTs') +
  theme(plot.title = element_text(hjust=0.5),
        axis.text.x = element_text(angle=45, hjust=1, vjust=1))
```


PLOT FOR MAIN MANUSCRIPT (Fig 4H)

```{r}
# png('Exp2_MeanRT.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(rtData1, model=='softmax(Q + gaze)'), aes(x=mean_RT, y=fit)) +
#   geom_abline(slope = 1, intercept=0, lty=3) +
#   geom_smooth(method='lm', col='maroon1') +
#   geom_point(pch=21, fill='grey', alpha=0.5) +
#   ylab('Mean RT (model)') +
#   xlab('Mean RT (data)') +
#   theme_classic(base_size = 11) +
#   theme(panel.grid.major.y = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3) +
#   ggtitle('Exp. 2: Learning')
# dev.off()
```


PLOT FOR SI APPENDIX (S9 Fig)

```{r}
# png('Exp2_IndividChoiceRT_All_Models.png', width=11.5, height=5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# grid.arrange(p1, p2, nrow=2, ncol=1)
# dev.off()
```


5. Relationships with model parameters (S4 Table)

```{r}
# merge with data
rtData <- rtData %>%
  left_join(params, by='subject')

# multiple regression predicting mean RTs from model parameters 
summary(lm(mean_RT ~ w_rel + learn_rate + Q_beta + gaze_beta + theta + upper_bound + threshold + t0, data=rtData))
```


## Response times for correct / incorrect choices

```{r}
# mean RTs for each participant by choice type
rtData2 <- dat %>%
  filter(block == 'learning', RT >= 250, RT <= 10000) %>%
  group_by(subject) %>%
  mutate(RT_rank = rank(RT) / n()) %>%
  ungroup() %>%
  group_by(subject, correct) %>%
  summarise(mean_RT = mean(RT),
            mean_rank = mean(RT_rank)) %>% 
  ungroup()
```


```{r}
with(rtData2, t.test(mean_RT[correct==0], mean_RT[correct==1], paired=T))
```

```{r}
with(rtData2, t.test(mean_rank[correct==0], mean_rank[correct==1], paired=T))
```



## Pooled response time distributions

1. Multiply incorrect choice RTs by -1 for plotting RT distributions.

```{r}
rtData_pooled <- dat %>%
  filter(RT >= 250, RT <= 10000) %>%
  mutate(RT_ = ifelse(correct==1, RT, -RT))
```


2. Do the same for simulated RTs after pooling across simulated participants.

```{r}
sims1_pooled <- abind::abind(sims1, along=1)
sims1_pooled <- apply(sims1_pooled, 2, c)
sims1_pooled <- data.frame(sims1_pooled)
sims1_pooled <- sims1_pooled %>% mutate(RT_ = ifelse(choice==1, RT, -RT))

sims2_pooled <- abind::abind(sims2, along=1)
sims2_pooled <- apply(sims2_pooled, 2, c)
sims2_pooled <- data.frame(sims2_pooled)
sims2_pooled <- sims2_pooled %>% mutate(RT_ = ifelse(choice==1, RT, -RT))

sims3_pooled <- abind::abind(sims3, along=1)
sims3_pooled <- apply(sims3_pooled, 2, c)
sims3_pooled <- data.frame(sims3_pooled)
sims3_pooled <- sims3_pooled %>% mutate(RT_ = ifelse(choice==1, RT, -RT))

sims4_pooled <- abind::abind(sims4, along=1)
sims4_pooled <- apply(sims4_pooled, 2, c)
sims4_pooled <- data.frame(sims4_pooled)
sims4_pooled <- sims4_pooled %>% mutate(RT_ = ifelse(X1==1, X2, -X2))

sims5_pooled <- abind::abind(sims5, along=1)
sims5_pooled <- apply(sims5_pooled, 2, c)
sims5_pooled <- data.frame(sims5_pooled)
sims5_pooled <- sims5_pooled %>% mutate(RT_ = ifelse(X1==1, X2, -X2))

sims6_pooled <- abind::abind(sims6, along=1)
sims6_pooled <- apply(sims6_pooled, 2, c)
sims6_pooled <- data.frame(sims6_pooled)
sims6_pooled <- sims6_pooled %>% mutate(RT_ = ifelse(X1==1, X2, -X2))

sims7_pooled <- abind::abind(sims7, along=1)
sims7_pooled <- apply(sims7_pooled, 2, c)
sims7_pooled <- data.frame(sims7_pooled)
sims7_pooled <- sims7_pooled %>% mutate(RT_ = ifelse(X1==1, X2, -X2))

sims8_pooled <- abind::abind(sims8, along=1)
sims8_pooled <- apply(sims8_pooled, 2, c)
sims8_pooled <- data.frame(sims8_pooled)
sims8_pooled <- sims8_pooled %>% mutate(RT_ = ifelse(X1==1, X2, -X2))
```


3. Plot correct and incorrect RT distributions with model simulations overlaid (S11 Fig)

```{r}
# png('Exp2_RTDists_All_Models.png', width=9, height=5, units='in', res=300)
# par(mfrow=c(2,4), mar=c(4.5,2,1.5,1))
# hist(rtData_pooled$RT_, nclass=50, freq=F, yaxt='n', ylab='', xlab='', border='grey50', col='grey', yaxs='i',
#      main='Q model', cex.main=0.9, font.main=2)
# lines(density(sims1_pooled$RT_), col='maroon1', lwd=2, yaxt='n')
# hist(rtData_pooled$RT_, nclass=50, freq=F, yaxt='n', ylab='', xlab='', border='grey50', col='grey', yaxs='i',
#      main='Q * gaze model', cex.main=0.9, font.main=2)
# lines(density(sims2_pooled$RT_), col='maroon1', lwd=2, yaxt='n')
# hist(rtData_pooled$RT_, nclass=50, freq=F, yaxt='n', ylab='', xlab='', border='grey50', col='grey', yaxs='i',
#      main='Q + gaze model', cex.main=0.9, font.main=2)
# lines(density(sims3_pooled$RT_), col='maroon1', lwd=2, yaxt='n')
# hist(rtData_pooled$RT_, nclass=50, freq=F, yaxt='n', ylab='', xlab='', border='grey50', col='grey', yaxs='i',
#      main='softmax(Q) model', cex.main=0.9, font.main=2)
# lines(density(sims4_pooled$RT_), col='maroon1', lwd=2, yaxt='n')
# hist(rtData_pooled$RT_, nclass=50, freq=F, yaxt='n', ylab='', xlab='', border='grey50', col='grey', yaxs='i',
#      main='softmax(Q*gaze) model', cex.main=0.9, font.main=2)
# lines(density(sims5_pooled$RT_), col='maroon1', lwd=2, yaxt='n')
# hist(rtData_pooled$RT_, nclass=50, freq=F, yaxt='n', ylab='', xlab='', border='grey50', col='grey', yaxs='i',
#      main='softmax(Q)*gaze model', cex.main=0.9, font.main=2)
# lines(density(sims6_pooled$RT_), col='maroon1', lwd=2, yaxt='n')
# hist(rtData_pooled$RT_, nclass=50, freq=F, yaxt='n', ylab='', xlab='', border='grey50', col='grey', yaxs='i',
#      main='softmax(Q) + gaze model', cex.main=0.9, font.main=2)
# lines(density(sims7_pooled$RT_), col='maroon1', lwd=2, yaxt='n')
# hist(rtData_pooled$RT_, nclass=50, freq=F, yaxt='n', ylab='', xlab='', border='grey50', col='grey', yaxs='i',
#      main='softmax(Q + gaze) model', cex.main=0.9, font.main=2)
# lines(density(sims8_pooled$RT_), col='maroon1', lwd=2, yaxt='n')
# dev.off()
```

4. Clean up

```{r}
rm(sims1_pooled, sims2_pooled, sims3_pooled, sims4_pooled, sims5_pooled,
   sims6_pooled, sims7_pooled)
gc()
```


## Choice accuracy as a function of relative gaze to the correct option: Learning phase


1. Start by computing each subject's choice accuracy for each of the five gaze buckets (i.e., quintiles), averaging across all learning trials.

```{r}
accByGazeData <- dat %>%
  filter(block == 'learning') %>%
  group_by(subject, gaze_bucket) %>%
  summarise(acc = mean(correct),
            n = n()) %>%
  ungroup()
```


2. Now do the same for the models.

```{r}
# Step 1: For each subject, compute accuracy within each gaze bucket for each of the 100 simulation runs (learning trials only)
# Returns a 5 (gaze buckets) x 100 (runs) x 83 (subjects) array
model1_accByGaze <- sapply(1:length(sims1), function(i) apply(sims1[[i]][1:120,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model2_accByGaze <- sapply(1:length(sims2), function(i) apply(sims2[[i]][1:120,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model3_accByGaze <- sapply(1:length(sims3), function(i) apply(sims3[[i]][1:120,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model4_accByGaze <- sapply(1:length(sims4), function(i) apply(sims4[[i]][1:120,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model5_accByGaze <- sapply(1:length(sims5), function(i) apply(sims5[[i]][1:120,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model6_accByGaze <- sapply(1:length(sims6), function(i) apply(sims6[[i]][1:120,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model7_accByGaze <- sapply(1:length(sims7), function(i) apply(sims7[[i]][1:120,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model8_accByGaze <- sapply(1:length(sims8), function(i) apply(sims8[[i]][1:120,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')


# Step 2: Average across simulation runs
model1_accByGaze <- apply(model1_accByGaze, c(1,3), FUN=mean)
model2_accByGaze <- apply(model2_accByGaze, c(1,3), FUN=mean)
model3_accByGaze <- apply(model3_accByGaze, c(1,3), FUN=mean)
model4_accByGaze <- apply(model4_accByGaze, c(1,3), FUN=mean)
model5_accByGaze <- apply(model5_accByGaze, c(1,3), FUN=mean)
model6_accByGaze <- apply(model6_accByGaze, c(1,3), FUN=mean)
model7_accByGaze <- apply(model7_accByGaze, c(1,3), FUN=mean)
model8_accByGaze <- apply(model8_accByGaze, c(1,3), FUN=mean)
```



3. Merge the model data with empirical data

```{r}
# Step 1: flatten model simulations to a vector (gaze buckets cycling faster than subjects)
model1_accByGaze <- c(model1_accByGaze)
model2_accByGaze <- c(model2_accByGaze)
model3_accByGaze <- c(model3_accByGaze)
model4_accByGaze <- c(model4_accByGaze)
model5_accByGaze <- c(model5_accByGaze)
model6_accByGaze <- c(model6_accByGaze)
model7_accByGaze <- c(model7_accByGaze)
model8_accByGaze <- c(model8_accByGaze)

# Step 2: create data frame 
modelAccByGaze <- data.frame(subject = rep(IDs, each=5),
                             gaze_bucket = rep(1:5, times=length(IDs)),
                             model_1 = model1_accByGaze,
                             model_2 = model2_accByGaze,
                             model_3 = model3_accByGaze,
                             model_4 = model4_accByGaze,
                             model_5 = model5_accByGaze,
                             model_6 = model6_accByGaze,
                             model_7 = model7_accByGaze,
                             model_8 = model8_accByGaze)

# Step 3: merge with empirical data
accByGazeData <- accByGazeData %>%
  left_join(modelAccByGaze, by=c('subject', 'gaze_bucket'))

# Step 4: reshape to long format
accByGazeData1 <- accByGazeData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# Step 5: turn model into a factor to control ordering
accByGazeData1 <- accByGazeData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))
```


4. Average across subjects


```{r}
accByGazeData2 <- accByGazeData1 %>%
  group_by(gaze_bucket, model) %>%
  summarise(mean_acc = mean(acc),
            sd_acc = sd(acc),
            n = n(),
            mean_fit = mean(fit)) %>%
  ungroup() %>%
  mutate(se = sd_acc / sqrt(n),
         upper = mean_acc + se,
         lower = mean_acc - se) 
```


WRITE DATA TO FILE

```{r}
#write.csv(accByGazeData2, 'Exp2_LearningAccByGazeQuintile_Fits.csv', row.names=F)
```


8. Plot mean and predicted choice accuracy across gaze buckets.

```{r}
p1 <- ggplot(accByGazeData2, aes(x=gaze_bucket, y=mean_acc)) +
  facet_wrap( ~model, nrow=1) +
  geom_bar(stat='identity', fill='grey70') +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
  geom_line(aes(y=mean_fit), col='maroon1', lty=1) +
  geom_point(aes(y=mean_fit), col='maroon1', pch=1) +
  ylab('Mean choice accuracy') +
  xlab('Excess relative gaze on the correct option (quintiles)') +
  ggtitle('Exp. 2 Learning Phase: Gaze Effects on Accuracy') +
  scale_y_continuous(limits=c(0,1)) +
  ggpubr::theme_pubclean() +
  theme(plot.title = element_text(hjust=0.5))
```

PLOT FOR MAIN MANUSCRIPT (Fig 5E)

```{r}
# png('Exp2_LearningAccByGazeQuintile.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(accByGazeData2, model %in% c('softmax(Q)', 'softmax(Q + gaze)')),
#        aes(x=gaze_bucket, y=mean_acc)) +
#   geom_line() +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
#   geom_point(stat='identity', fill='grey70', pch=21) +
#   geom_line(aes(y=mean_fit, col=model), lty=2) +
#   geom_point(aes(y=mean_fit, col=model), pch=1, show.legend=FALSE) +
#   ylab('Mean choice accuracy') +
#   xlab('Excess prop. gaze (quintiles)') +
#   scale_y_continuous(limits=c(0,1)) +
#   theme_classic(base_size = 11) +
#   theme(legend.position = 'inside',
#         legend.position.inside = c(.65, .2),
#         legend.title = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   scale_color_manual(values=c('deepskyblue', 'maroon1')) +
#   ggtitle('Exp. 2: Learning')
# dev.off()
```



## Choice response time as a function of relative gaze to the correct option: Learning phase


1. Start by computing each subject's mean RT for each of the five gaze buckets (quintiles), averaging across all learning trials.

```{r}
RTByGazeData <- dat %>%
  filter(block == 'learning', RT >= 250, RT <= 10000) %>%
  group_by(subject, gaze_bucket) %>%
  summarise(mean_RT = mean(RT),
            n = n()) %>%
  ungroup()
```


2. Now do the same for the models.

```{r}
# Step 1: For each subject, compute mean RT within each gaze bucket for each of the 100 simulation runs (learning trials only)
# Returns a 5 (gaze buckets) x 100 (runs) x 83 (subjects) array
model1_RTByGaze <- sapply(1:length(sims1), function(i) apply(sims1[[i]][1:120,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model2_RTByGaze <- sapply(1:length(sims2), function(i) apply(sims2[[i]][1:120,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model3_RTByGaze <- sapply(1:length(sims3), function(i) apply(sims3[[i]][1:120,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model4_RTByGaze <- sapply(1:length(sims4), function(i) apply(sims4[[i]][1:120,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model5_RTByGaze <- sapply(1:length(sims5), function(i) apply(sims5[[i]][1:120,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model6_RTByGaze <- sapply(1:length(sims6), function(i) apply(sims6[[i]][1:120,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model7_RTByGaze <- sapply(1:length(sims7), function(i) apply(sims7[[i]][1:120,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')
model8_RTByGaze <- sapply(1:length(sims8), function(i) apply(sims8[[i]][1:120,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[1:120]), FUN=mean)[,2]}), simplify='array')


# Step 2: Average across simulation runs
model1_RTByGaze <- apply(model1_RTByGaze, c(1,3), FUN=mean)
model2_RTByGaze <- apply(model2_RTByGaze, c(1,3), FUN=mean)
model3_RTByGaze <- apply(model3_RTByGaze, c(1,3), FUN=mean)
model4_RTByGaze <- apply(model4_RTByGaze, c(1,3), FUN=mean)
model5_RTByGaze <- apply(model5_RTByGaze, c(1,3), FUN=mean)
model6_RTByGaze <- apply(model6_RTByGaze, c(1,3), FUN=mean)
model7_RTByGaze <- apply(model7_RTByGaze, c(1,3), FUN=mean)
model8_RTByGaze <- apply(model8_RTByGaze, c(1,3), FUN=mean)
```


3. Merge the model data with empirical data

```{r}
# Step 1: flatten model simulations to a vector (gaze buckets cycling faster than subjects)
model1_RTByGaze <- c(model1_RTByGaze)
model2_RTByGaze <- c(model2_RTByGaze)
model3_RTByGaze <- c(model3_RTByGaze)
model4_RTByGaze <- c(model4_RTByGaze)
model5_RTByGaze <- c(model5_RTByGaze)
model6_RTByGaze <- c(model6_RTByGaze)
model7_RTByGaze <- c(model7_RTByGaze)
model8_RTByGaze <- c(model8_RTByGaze)

# Step 2: create data frame 
modelRTByGaze <- data.frame(subject = rep(IDs, each=5),
                             gaze_bucket = rep(1:5, times=length(IDs)),
                             model_1 = model1_RTByGaze,
                             model_2 = model2_RTByGaze,
                             model_3 = model3_RTByGaze,
                             model_4 = model4_RTByGaze,
                             model_5 = model5_RTByGaze,
                             model_6 = model6_RTByGaze,
                             model_7 = model7_RTByGaze,
                             model_8 = model8_RTByGaze)

# Step 3: merge with empirical data
RTByGazeData <- RTByGazeData %>%
  left_join(modelRTByGaze, by=c('subject', 'gaze_bucket'))

# Step 4: reshape to long format
RTByGazeData1 <- RTByGazeData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# Step 5: turn model into a factor to control ordering
RTByGazeData1 <- RTByGazeData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))
```


4. Average across subjects


```{r}
RTByGazeData2 <- RTByGazeData1 %>%
  group_by(gaze_bucket, model) %>%
  summarise(avg_mean_RT = mean(mean_RT),
            sd_mean_RT = sd(mean_RT),
            n = n(),
            mean_fit = mean(fit)) %>%
  ungroup() %>%
  mutate(se = sd_mean_RT / sqrt(n),
         upper = avg_mean_RT + se,
         lower = avg_mean_RT - se) 
```

WRITE DATA TO FILE

```{r}
#write.csv(RTByGazeData2, 'Exp2_LearningMeanRTByGazeQuintile_Fits.csv', row.names=F)
```


5. Plot mean and predicted choice RTs across gaze buckets.

```{r}
p2 <- ggplot(RTByGazeData2, aes(x=gaze_bucket, y=avg_mean_RT)) +
  facet_wrap( ~model, nrow=1) +
  geom_bar(stat='identity', fill='grey70') +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
  geom_line(aes(y=mean_fit), col='maroon1', lty=1) +
  geom_point(aes(y=mean_fit), col='maroon1', pch=1) +
  ylab('Mean RT (ms)') +
  xlab('Excess relative gaze on the correct option (quintiles)') +
  scale_y_continuous(limits=c(0,2000)) +
  ggpubr::theme_pubclean() +
  ggtitle('Exp. 2 Learning Phase: Gaze Effects on Mean RT') +
  theme(plot.title = element_text(hjust=0.5))
```

PLOT FOR MAIN MANUSCRIPT (Fig 5F)

```{r}
# png('Exp2_LearningMeanRTByGazeQuintile.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(RTByGazeData2, model %in% c('softmax(Q)', 'softmax(Q + gaze)')),
#        aes(x=gaze_bucket, y=avg_mean_RT)) +
#   geom_line() +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
#   geom_point(stat='identity', fill='grey70', pch=21) +
#   geom_line(aes(y=mean_fit, col=model), lty=2, show.legend=FALSE) +
#   geom_point(aes(y=mean_fit, col=model), pch=1, show.legend=FALSE) +
#   ylab('Mean RT (ms)') +
#   xlab('Excess prop. gaze (quintiles)') +
#   scale_y_continuous(limits=c(500, 2000)) +
#   theme_classic(base_size = 11) +
#   theme(legend.position = 'inside',
#         legend.position.inside = c(.65, .2),
#         legend.title = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   scale_color_manual(values=c('deepskyblue', 'maroon1')) +
#   ggtitle('Exp. 2: Learning')
# dev.off()
```


PLOT FOR SI APPENDIX (S14 Fig)

```{r}
# png('Exp2_GazeEffects_All_Models.png', width=11.5, height=5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# grid.arrange(p1, p2, nrow=2, ncol=1)
# dev.off()
```

6. Load model fits and extract the estimated gaze parameters

```{r}
model8_fits <- readRDS('../modeling/results/free-theta/model8_fits.RDS')

params <- sapply(model8_fits, function(X) X$optim$bestmem)

gaze_params <- data.frame(subject=IDs, gaze_beta=params[4,], theta=params[8,])

```


7. Merge parameter data with the individual accuracy x gaze quintile data

```{r}
accByGazeData1 <- accByGazeData1 %>%
  left_join(gaze_params, by='subject')
```


8. Gaze effect on accuracy: (Acc. in 5th gaze quintile) - (Acc. in 1st gaze quintile)

```{r}
gazeEffectData <- accByGazeData1 %>%
  mutate(contrast_weight = case_when(gaze_bucket == 1 ~ -1,
                                     gaze_bucket == 5 ~ 1,
                                     .default=0)) %>%
  group_by(subject, gaze_beta, theta, model) %>%
  summarise(gaze_effect = sum(contrast_weight * acc),
            fitted_effect = sum(contrast_weight * fit)) %>%
  ungroup()
```


WRITE DATA TO FILE

```{r}
#write.csv(gazeEffectData, 'Exp2_LearningGazeEffect_Fits.csv', row.names=F)
```


9. Plot of empirical vs. fitted gaze effects

```{r}
# ggplot(gazeEffectData, aes(x=gaze_effect, y=fitted_effect)) +
#   facet_wrap( ~model, nrow=2) +
#   geom_abline(slope = 1, intercept=0, lty=3) +
#   geom_smooth(method='lm', col='green') +
#   geom_point(pch=21, fill='grey', alpha=0.5) +
#   ylab('Gaze effect on accuracy (model)') +
#   xlab('Gaze effect on accuracy (data) ') +
#   ggpubr::theme_pubclean() +
#   theme(panel.grid.major.y = element_blank()) +
#   ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3)
```


PLOT FOR MAIN MANUSCRIPT (Fig 5G)

```{r}
# png('Exp2_LearningGazeEffects.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(gazeEffectData, model=='softmax(Q + gaze)'), aes(x=gaze_effect, y=fitted_effect)) +
#   geom_abline(slope = 1, intercept=0, lty=3) +
#   geom_smooth(method='lm', col='maroon1') +
#   geom_point(pch=21, fill='grey', alpha=0.5) +
#   ylab(expression(paste(Delta, 'accuracy'['Q5 - Q1'], ' (model)'))) +
#   xlab(expression(paste(Delta, 'accuracy'['Q5 - Q1'], ' (data)'))) +
#   theme_classic(base_size = 11) +
#   theme(panel.grid.major.y = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3) +
#   ggtitle('Exp. 2: Learning')
# dev.off()
```


PLOT FOR MAIN MANUSCRIPT (Fig 5H)

```{r}
# png('Exp2_LearningGazeEffects2.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(gazeEffectData, model=='softmax(Q + gaze)'), aes(x=gaze_effect, y=gaze_beta*theta)) +
#   geom_smooth(method='lm', col='maroon1') +
#   geom_point(pch=21, fill='grey', alpha=0.5) +
#   ylab(expression(theta * beta['gaze'])) +
#   xlab(expression(paste(Delta, 'accuracy'['Q5 - Q1'], ' (data)'))) +
#   theme_classic(base_size = 11) +
#   theme(panel.grid.major.y = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3) +
#   ggtitle('Exp. 2: Learning')
# dev.off()
```


## Transfer test analysis

Categorize trials as congruent, neutral (++ / --), or incongruent.

```{r}
transferData <- dat %>% filter(block == 'transfer')

transferData <- transferData %>%
  mutate(trial_type = case_when((correct_resp == 'left') & 
                                  (left_index %% 2 == 0) & 
                                  (right_index %% 2 == 1) ~ 'C',
                                (correct_resp == 'left') & 
                                  (left_index %% 2 == 1) & 
                                  (right_index %% 2 == 0) ~ 'I',
                                (correct_resp == 'right') & 
                                  (left_index %% 2 == 1) & 
                                  (right_index %% 2 == 0) ~ 'C',
                                (correct_resp == 'right') & 
                                  (left_index %% 2 == 0) & 
                                  (right_index %% 2 == 1) ~ 'I',
                                (left_index %% 2 == 0) & 
                                  (right_index %% 2 == 0) ~ 'N++',
                                (left_index %% 2 == 1) & 
                                  (right_index %% 2 == 1) ~ 'N--'),
         trial_type = factor(trial_type, levels=c('C', 'N++', 'N--', 'I')))
```


### Choice Accuracy by Trial Type

1. Get individual subject accuracies for each trial type, and copy to the list of individual datasets (to facilitate the next step).


```{r}
transferAccData <- transferData %>%
  group_by(subject, trial_type) %>%
  summarise(acc = mean(correct)) %>%
  ungroup()


# copy trial type variable to the list of datasets
for (i in 1:length(IDs)) {
  datasets[[i]]$trial_type <- transferData$trial_type[transferData$subject == IDs[i]]
}

```


2. Now for each subject, get the model-simulated accuracies for each trial type.

```{r}
# Step 1: For each subject, compute accuracy within each trial type for each of the 100 simulation runs (transfer trials only)
# Returns a 5 (gaze buckets) x 100 (runs) x 83 (subjects) array
model1_accByTType <- sapply(1:length(sims1), function(i) apply(sims1[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model2_accByTType <- sapply(1:length(sims2), function(i) apply(sims2[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model3_accByTType <- sapply(1:length(sims3), function(i) apply(sims3[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model4_accByTType <- sapply(1:length(sims4), function(i) apply(sims4[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model5_accByTType <- sapply(1:length(sims5), function(i) apply(sims5[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model6_accByTType <- sapply(1:length(sims6), function(i) apply(sims6[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model7_accByTType <- sapply(1:length(sims7), function(i) apply(sims7[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model8_accByTType <- sapply(1:length(sims8), function(i) apply(sims8[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')


# Step 2: Average across simulation runs
model1_accByTType <- apply(model1_accByTType, c(1,3), FUN=mean)
model2_accByTType <- apply(model2_accByTType, c(1,3), FUN=mean)
model3_accByTType <- apply(model3_accByTType, c(1,3), FUN=mean)
model4_accByTType <- apply(model4_accByTType, c(1,3), FUN=mean)
model5_accByTType <- apply(model5_accByTType, c(1,3), FUN=mean)
model6_accByTType <- apply(model6_accByTType, c(1,3), FUN=mean)
model7_accByTType <- apply(model7_accByTType, c(1,3), FUN=mean)
model8_accByTType <- apply(model8_accByTType, c(1,3), FUN=mean)
```



3. Merge the model data with empirical data

```{r}
# Step 1: flatten model simulations to a vector (trial_types cycling faster than subjects)
model1_accByTType <- c(model1_accByTType)
model2_accByTType <- c(model2_accByTType)
model3_accByTType <- c(model3_accByTType)
model4_accByTType <- c(model4_accByTType)
model5_accByTType <- c(model5_accByTType)
model6_accByTType <- c(model6_accByTType)
model7_accByTType <- c(model7_accByTType)
model8_accByTType <- c(model8_accByTType)

# Step 2: create data frame 
modelAccByTType <- data.frame(subject = rep(IDs, each=4),
                              trial_type = factor(rep(levels(transferData$trial_type), times=length(IDs)),
                                                  levels=levels(transferData$trial_type)),
                              model_1 = model1_accByTType,
                              model_2 = model2_accByTType,
                              model_3 = model3_accByTType,
                              model_4 = model4_accByTType,
                              model_5 = model5_accByTType,
                              model_6 = model6_accByTType,
                              model_7 = model7_accByTType,
                              model_8 = model8_accByTType)

# Step 3: merge with empirical data
transferAccData <- transferAccData %>%
  left_join(modelAccByTType, by=c('subject', 'trial_type'))

# Step 4: reshape to long format
transferAccData1 <- transferAccData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# Step 5: turn model into a factor to control ordering
transferAccData1 <- transferAccData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))
```


4. Average across subjects


```{r}
transferAccData2 <- transferAccData1 %>%
  group_by(trial_type, model) %>%
  summarise(mean_acc = mean(acc),
            sd_acc = sd(acc),
            n = n(),
            mean_fit = mean(fit)) %>%
  ungroup() %>%
  mutate(se = sd_acc / sqrt(n),
         upper = mean_acc + se,
         lower = mean_acc - se) 
```

WRITE DATA TO FILE

```{r}
#write.csv(transferAccData2, 'Exp2_TransferAccByTrialType_Fits.csv', row.names=F)
```


5. Plot mean and predicted choice accuracy across trial types.

```{r}
p1 <- ggplot(transferAccData2, aes(x=trial_type, y=mean_acc)) +
  facet_wrap( ~model, nrow=1) +
  geom_bar(stat='identity', fill='grey70') +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
  geom_line(aes(y=mean_fit, group=1), col='maroon1', lty=1) +
  geom_point(aes(y=mean_fit), col='maroon1', pch=1) +
  ylab('Mean choice accuracy') +
  xlab('Trial Type') +
  ggtitle('Exp. 2 Transfer Test: Mean Accuracy by Trial Type') +
  scale_y_continuous(limits=c(0,1)) +
  ggpubr::theme_pubclean() +
  theme(plot.title = element_text(hjust=0.5))
```

PLOT FOR MAIN MANUSCRIPT (Fig 4I)

```{r}
# png('Exp2_TransferAccByTrialType.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(transferAccData2, model %in% c('Q + gaze', 'softmax(Q + gaze)')),
#        aes(x=trial_type, y=mean_acc)) +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
#   geom_point(stat='identity', fill='grey70', pch=21) +
#   geom_line(aes(y=mean_fit, col=model, group=model), lty=2) +
#   geom_point(aes(y=mean_fit, col=model), pch=1, show.legend=FALSE) +
#   ylab('Mean choice accuracy') +
#   xlab('Trial Type') +
#   scale_y_continuous(limits=c(0,1)) +
#   theme_classic(base_size = 11) +
#   theme(legend.position = 'inside',
#         legend.position.inside = c(.42, .2),
#         legend.title = element_blank(),
#         legend.background = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   scale_color_manual(values=c('green', 'maroon1')) +
#   ggtitle('Exp. 2: Transfer')
# dev.off()
```


### Mean RT by Trial Type

1. Get individual subject mean RTs for each trial type.


```{r}
transferRTData <- transferData %>%
  group_by(subject, trial_type) %>%
  summarise(mean_RT = mean(RT)) %>%
  ungroup()
```


2. Now for each subject, get the model-simulated mean RTs for each trial type 
** Note: assumes trial type variable has been added to list of data sets (see above) **

```{r}
# Step 1: For each subject, compute mean RT within each trial type for each of the 100 simulation runs (transfer trials only)
# Returns a 5 (gaze buckets) x 100 (runs) x 83 (subjects) array
model1_RTByTType <- sapply(1:length(sims1), function(i) apply(sims1[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model2_RTByTType <- sapply(1:length(sims2), function(i) apply(sims2[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model3_RTByTType <- sapply(1:length(sims3), function(i) apply(sims3[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model4_RTByTType <- sapply(1:length(sims4), function(i) apply(sims4[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model5_RTByTType <- sapply(1:length(sims5), function(i) apply(sims5[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model6_RTByTType <- sapply(1:length(sims6), function(i) apply(sims6[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model7_RTByTType <- sapply(1:length(sims7), function(i) apply(sims7[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')
model8_RTByTType <- sapply(1:length(sims8), function(i) apply(sims8[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ datasets[[i]]$trial_type, FUN=mean)[,2]}), simplify='array')


# Step 2: Average across simulation runs
model1_RTByTType <- apply(model1_RTByTType, c(1,3), FUN=mean)
model2_RTByTType <- apply(model2_RTByTType, c(1,3), FUN=mean)
model3_RTByTType <- apply(model3_RTByTType, c(1,3), FUN=mean)
model4_RTByTType <- apply(model4_RTByTType, c(1,3), FUN=mean)
model5_RTByTType <- apply(model5_RTByTType, c(1,3), FUN=mean)
model6_RTByTType <- apply(model6_RTByTType, c(1,3), FUN=mean)
model7_RTByTType <- apply(model7_RTByTType, c(1,3), FUN=mean)
model8_RTByTType <- apply(model8_RTByTType, c(1,3), FUN=mean)
```



3. Merge the model data with empirical data

```{r}
# Step 1: flatten model simulations to a vector (trial_types cycling faster than subjects)
model1_RTByTType <- c(model1_RTByTType)
model2_RTByTType <- c(model2_RTByTType)
model3_RTByTType <- c(model3_RTByTType)
model4_RTByTType <- c(model4_RTByTType)
model5_RTByTType <- c(model5_RTByTType)
model6_RTByTType <- c(model6_RTByTType)
model7_RTByTType <- c(model7_RTByTType)
model8_RTByTType <- c(model8_RTByTType)

# Step 2: create data frame 
modelRTByTType <- data.frame(subject = rep(IDs, each=4),
                             trial_type = factor(rep(levels(transferData$trial_type), times=length(IDs)),
                                                 levels=levels(transferData$trial_type)),
                             model_1 = model1_RTByTType,
                             model_2 = model2_RTByTType,
                             model_3 = model3_RTByTType,
                             model_4 = model4_RTByTType,
                             model_5 = model5_RTByTType,
                             model_6 = model6_RTByTType,
                             model_7 = model7_RTByTType,
                             model_8 = model8_RTByTType)

# Step 3: merge with empirical data
transferRTData <- transferRTData %>%
  left_join(modelRTByTType, by=c('subject', 'trial_type'))

# Step 4: reshape to long format
transferRTData1 <- transferRTData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# Step 5: turn model into a factor to control ordering
transferRTData1 <- transferRTData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))
```


4. Average across subjects


```{r}
transferRTData2 <- transferRTData1 %>%
  group_by(trial_type, model) %>%
  summarise(avg_mean_RT = mean(mean_RT),
            sd_mean_RT = sd(mean_RT),
            n = n(),
            mean_fit = mean(fit)) %>%
  ungroup() %>%
  mutate(se = sd_mean_RT / sqrt(n),
         upper = avg_mean_RT + se,
         lower = avg_mean_RT - se) 
```


WRITE DATA TO FILE

```{r}
#write.csv(transferRTData2, 'Exp2_TransferMeanRTByTrialType_Fits.csv', row.names=F)
```


5. Plot mean and predicted RT across trial types.

```{r}
p2 <- ggplot(transferRTData2, aes(x=trial_type, y=avg_mean_RT)) +
  facet_wrap( ~model, nrow=1) +
  geom_bar(stat='identity', fill='grey70') +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
  geom_line(aes(y=mean_fit, group=1), col='maroon1', lty=1) +
  geom_point(aes(y=mean_fit), col='maroon1', pch=1) +
  ylab('Mean RT (ms)') +
  xlab('Trial Type') +
  ggtitle('Exp. 2 Transfer Test: Mean RT by Trial Type') +
  #scale_y_continuous(limits=c(0,1)) +
  ggpubr::theme_pubclean() +
  theme(plot.title = element_text(hjust=0.5))
```


PLOT FOR MAIN MANUSCRIPT (Fig 4J)

```{r}
# png('Exp2_TransferMeanRTByTrialType.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(transferRTData2, model %in% c('Q + gaze', 'softmax(Q + gaze)')),
#        aes(x=trial_type, y=avg_mean_RT)) +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
#   geom_point(stat='identity', fill='grey70', pch=21) +
#   geom_line(aes(y=mean_fit, col=model, group=model), lty=2, show.legend=FALSE) +
#   geom_point(aes(y=mean_fit, col=model), pch=1, show.legend=FALSE) +
#   ylab('Mean RT (ms)') +
#   xlab('Trial Type') +
#   scale_y_continuous(limits=c(500, 2000)) +
#   theme_classic(base_size = 11) +
#   theme(legend.position = 'inside',
#         legend.position.inside = c(.42, .2),
#         legend.title = element_blank(),
#         legend.background = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   scale_color_manual(values=c('green', 'maroon1')) +
#   ggtitle('Exp. 2: Transfer')
# dev.off()
```


PLOT FOR SI APPENDIX (S12 Fig)

```{r}
# png('Exp2_TransferTest_All_Models.png', width=11.5, height=5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# grid.arrange(p1, p2, nrow=2, ncol=1)
# dev.off()
```


### Incongruent trial accuracies

1. Extract relative encoding parameter and merge with the individual transfer accuracy data

```{r}
# load model 8 fits
model8_fits <- readRDS('../modeling/results/free-theta/model8_fits.RDS')

# extract parameter estimates
params <- data.frame(t(sapply(model8_fits, function(X) X$optim$bestmem)))
names(params) <- c('t0','learn_rate','Q_beta','gaze_beta','threshold_sep','upper_bound','w_rel','theta')
params$threshold <- params$upper_bound + params$threshold_sep

# add subject ID column
params$subject <- IDs

transferAccData1 <- transferAccData1 %>%
  left_join(params, by='subject')
```


2. Incongruent trials only

```{r}
incongTrialData <- transferAccData1 %>% filter(trial_type == 'I')
```


WRITE DATA TO FILE

```{r}
#write.csv(incongTrialData, 'Exp2_TransferIncongTrialAcc_Fits.csv', row.names=F)
```


3. Plot of empirical vs. fitted accuracies on incongruent transfer trials

```{r}
# ggplot(incongTrialData, aes(x=acc, y=fit)) +
#   facet_wrap( ~model, nrow=2) +
#   geom_abline(slope = 1, intercept=0, lty=3) +
#   geom_smooth(method='lm', col='maroon1') +
#   geom_point(pch=21, fill='grey', alpha=0.5) +
#   ylab('Incong. accuracy (model)') +
#   xlab('Incong. accuracy (data) ') +
#   ggpubr::theme_pubclean() +
#   theme(panel.grid.major.y = element_blank()) +
#   ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3)
```

PLOT FOR MAIN MANUSCRIPT (Fig 4K)

```{r}
# png('Exp2_TransferIncongTrialAcc.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(incongTrialData, model=='softmax(Q + gaze)'), aes(x=acc, y=fit)) +
#   geom_abline(slope = 1, intercept=0, lty=3) +
#   geom_smooth(method='lm', col='maroon1') +
#   geom_point(pch=21, fill='grey', alpha=0.5) +
#   ylab('Incong. accuracy (model)') +
#   xlab('Incong. accuracy (data)') +
#   theme_classic(base_size = 11) +
#   theme(panel.grid.major.y = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3) +
#   ggtitle('Exp. 2: Transfer')
# dev.off()
```


PLOT FOR MAIN MANUSCRIPT (Fig 4L)

```{r}
# png('Exp2_TransferIncongTrialAcc2.png', width=2.5, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(incongTrialData, model=='softmax(Q) + gaze'), aes(x=acc, y=w_rel)) +
#   #geom_abline(slope = 1, intercept=0, lty=3) +
#   geom_smooth(method='lm', col='maroon1') +
#   geom_point(pch=21, fill='grey', alpha=0.5) +
#   ylab(expression('w'['rel'])) +
#   xlab('Incong. accuracy (data)') +
#   #scale_y_continuous(limits=c(0,1)) +
#   theme_classic(base_size = 11) +
#   theme(panel.grid.major.y = element_blank(),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   ggpubr::stat_cor(aes(label = ..rr.label..), method='pearson', cex=3, label.x.npc=.75) +
#   ggtitle('Exp. 2: Transfer')
# dev.off()
```

Table S5

```{r}
# multiple regression predicting incongruent trial accuracy from RL-SSM parameters
summary(lm(acc ~ w_rel + learn_rate + Q_beta + gaze_beta + upper_bound + threshold + t0 + theta,
           data = subset(incongTrialData, model == 'softmax(Q + gaze)')))
```



## Choice accuracy as a function of relative gaze to the correct option: Transfer test


1. Start by computing each subject's choice accuracy for each of the five gaze buckets, averaging across all transfer trials.

```{r}
accByGazeData <- dat %>%
  filter(block == 'transfer') %>%
  group_by(subject, gaze_bucket) %>%
  summarise(acc = mean(correct),
            n = n()) %>%
  ungroup()
```


2. Now do the same for the models.

```{r}
# Step 1: For each subject, compute accuracy within each gaze bucket for each of the 100 simulation runs (transfer trials only)
# Returns a 5 (gaze buckets) x 100 (runs) x 83 (subjects) array
model1_accByGaze <- sapply(1:length(sims1), function(i) apply(sims1[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model2_accByGaze <- sapply(1:length(sims2), function(i) apply(sims2[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model3_accByGaze <- sapply(1:length(sims3), function(i) apply(sims3[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model4_accByGaze <- sapply(1:length(sims4), function(i) apply(sims4[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model5_accByGaze <- sapply(1:length(sims5), function(i) apply(sims5[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model6_accByGaze <- sapply(1:length(sims6), function(i) apply(sims6[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model7_accByGaze <- sapply(1:length(sims7), function(i) apply(sims7[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model8_accByGaze <- sapply(1:length(sims8), function(i) apply(sims8[[i]][121:232,1,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')


# Step 2: Average across simulation runs
model1_accByGaze <- apply(model1_accByGaze, c(1,3), FUN=mean)
model2_accByGaze <- apply(model2_accByGaze, c(1,3), FUN=mean)
model3_accByGaze <- apply(model3_accByGaze, c(1,3), FUN=mean)
model4_accByGaze <- apply(model4_accByGaze, c(1,3), FUN=mean)
model5_accByGaze <- apply(model5_accByGaze, c(1,3), FUN=mean)
model6_accByGaze <- apply(model6_accByGaze, c(1,3), FUN=mean)
model7_accByGaze <- apply(model7_accByGaze, c(1,3), FUN=mean)
model8_accByGaze <- apply(model8_accByGaze, c(1,3), FUN=mean)
```



3. Merge the model data with empirical data

```{r}
# Step 1: flatten model simulations to a vector (gaze buckets cycling faster than subjects)
model1_accByGaze <- c(model1_accByGaze)
model2_accByGaze <- c(model2_accByGaze)
model3_accByGaze <- c(model3_accByGaze)
model4_accByGaze <- c(model4_accByGaze)
model5_accByGaze <- c(model5_accByGaze)
model6_accByGaze <- c(model6_accByGaze)
model7_accByGaze <- c(model7_accByGaze)
model8_accByGaze <- c(model8_accByGaze)

# Step 2: create data frame 
modelAccByGaze <- data.frame(subject = rep(IDs, each=5),
                             gaze_bucket = rep(1:5, times=length(IDs)),
                             model_1 = model1_accByGaze,
                             model_2 = model2_accByGaze,
                             model_3 = model3_accByGaze,
                             model_4 = model4_accByGaze,
                             model_5 = model5_accByGaze,
                             model_6 = model6_accByGaze,
                             model_7 = model7_accByGaze,
                             model_8 = model8_accByGaze)

# Step 3: merge with empirical data
accByGazeData <- accByGazeData %>%
  left_join(modelAccByGaze, by=c('subject', 'gaze_bucket'))

# Step 4: reshape to long format
accByGazeData1 <- accByGazeData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# Step 5: turn model into a factor to control ordering
accByGazeData1 <- accByGazeData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))
```


4. Average across subjects


```{r}
accByGazeData2 <- accByGazeData1 %>%
  group_by(gaze_bucket, model) %>%
  summarise(mean_acc = mean(acc),
            sd_acc = sd(acc),
            n = n(),
            mean_fit = mean(fit)) %>%
  ungroup() %>%
  mutate(se = sd_acc / sqrt(n),
         upper = mean_acc + se,
         lower = mean_acc - se) 
```


5. Plot mean and predicted choice accuracy across gaze buckets.

```{r}
p1 <- ggplot(accByGazeData2, aes(x=gaze_bucket, y=mean_acc)) +
  facet_wrap( ~model, nrow=1) +
  geom_bar(stat='identity', fill='grey70') +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
  geom_line(aes(y=mean_fit), col='maroon1', lty=1) +
  geom_point(aes(y=mean_fit), col='maroon1', pch=1) +
  ylab('Mean choice accuracy') +
  xlab('Excess relative gaze on the correct option (quintiles)') +
  ggtitle('Exp. 2 Transfer Test: Gaze Effects on Accuracy') +
  scale_y_continuous(limits=c(0,1)) +
  ggpubr::theme_pubclean() +
  theme(plot.title = element_text(hjust=0.5))
```


## Choice response time as a function of relative gaze to the correct option: Transfer test


1. Start by computing each subject's mean RT for each of the five gaze buckets, averaging across all transfer trials.

```{r}
RTByGazeData <- dat %>%
  filter(block == 'transfer', RT >= 250, RT <= 10000) %>%
  group_by(subject, gaze_bucket) %>%
  summarise(mean_RT = mean(RT),
            n = n()) %>%
  ungroup()
```


2. Now do the same for the models.

```{r}
# Step 1: For each subject, compute mean RT within each gaze bucket for each of the 100 simulation runs (transfer trials only)
# Returns a 5 (gaze buckets) x 100 (runs) x 83 (subjects) array
model1_RTByGaze <- sapply(1:length(sims1), function(i) apply(sims1[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model2_RTByGaze <- sapply(1:length(sims2), function(i) apply(sims2[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model3_RTByGaze <- sapply(1:length(sims3), function(i) apply(sims3[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model4_RTByGaze <- sapply(1:length(sims4), function(i) apply(sims4[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model5_RTByGaze <- sapply(1:length(sims5), function(i) apply(sims5[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model6_RTByGaze <- sapply(1:length(sims6), function(i) apply(sims6[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model7_RTByGaze <- sapply(1:length(sims7), function(i) apply(sims7[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')
model8_RTByGaze <- sapply(1:length(sims8), function(i) apply(sims8[[i]][121:232,2,], 2, FUN=function(x_j) {
  aggregate(x_j ~ factor(datasets[[i]]$gaze_bucket[121:232]), FUN=mean)[,2]}), simplify='array')


# Step 2: Average across simulation runs
model1_RTByGaze <- apply(model1_RTByGaze, c(1,3), FUN=mean)
model2_RTByGaze <- apply(model2_RTByGaze, c(1,3), FUN=mean)
model3_RTByGaze <- apply(model3_RTByGaze, c(1,3), FUN=mean)
model4_RTByGaze <- apply(model4_RTByGaze, c(1,3), FUN=mean)
model5_RTByGaze <- apply(model5_RTByGaze, c(1,3), FUN=mean)
model6_RTByGaze <- apply(model6_RTByGaze, c(1,3), FUN=mean)
model7_RTByGaze <- apply(model7_RTByGaze, c(1,3), FUN=mean)
model8_RTByGaze <- apply(model8_RTByGaze, c(1,3), FUN=mean)
```


3. Merge the model data with empirical data

```{r}
# Step 1: flatten model simulations to a vector (gaze buckets cycling faster than subjects)
model1_RTByGaze <- c(model1_RTByGaze)
model2_RTByGaze <- c(model2_RTByGaze)
model3_RTByGaze <- c(model3_RTByGaze)
model4_RTByGaze <- c(model4_RTByGaze)
model5_RTByGaze <- c(model5_RTByGaze)
model6_RTByGaze <- c(model6_RTByGaze)
model7_RTByGaze <- c(model7_RTByGaze)
model8_RTByGaze <- c(model8_RTByGaze)

# Step 2: create data frame 
modelRTByGaze <- data.frame(subject = rep(IDs, each=5),
                             gaze_bucket = rep(1:5, times=length(IDs)),
                             model_1 = model1_RTByGaze,
                             model_2 = model2_RTByGaze,
                             model_3 = model3_RTByGaze,
                             model_4 = model4_RTByGaze,
                             model_5 = model5_RTByGaze,
                             model_6 = model6_RTByGaze,
                             model_7 = model7_RTByGaze,
                             model_8 = model8_RTByGaze)

# Step 3: merge with empirical data
RTByGazeData <- RTByGazeData %>%
  left_join(modelRTByGaze, by=c('subject', 'gaze_bucket'))

# Step 4: reshape to long format
RTByGazeData1 <- RTByGazeData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='fit')

# Step 5: turn model into a factor to control ordering
RTByGazeData1 <- RTByGazeData1 %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels))
```


4. Average across subjects


```{r}
RTByGazeData2 <- RTByGazeData1 %>%
  group_by(gaze_bucket, model) %>%
  summarise(avg_mean_RT = mean(mean_RT),
            sd_mean_RT = sd(mean_RT),
            n = n(),
            mean_fit = mean(fit)) %>%
  ungroup() %>%
  mutate(se = sd_mean_RT / sqrt(n),
         upper = avg_mean_RT + se,
         lower = avg_mean_RT - se) 
```


5. Plot mean and predicted choice RTs across gaze buckets.

```{r}
p2 <- ggplot(RTByGazeData2, aes(x=gaze_bucket, y=avg_mean_RT)) +
  facet_wrap( ~model, nrow=1) +
  geom_bar(stat='identity', fill='grey70') +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
  geom_line(aes(y=mean_fit), col='maroon1', lty=1) +
  geom_point(aes(y=mean_fit), col='maroon1', pch=1) +
  ylab('Mean RT (ms)') +
  xlab('Excess relative gaze on the correct option (quintiles)') +
  ggtitle('Exp. 2 Transfer Test: Gaze Effects on Mean RT') +
  scale_y_continuous(limits=c(0,2000)) +
  ggpubr::theme_pubclean() +
  theme(plot.title = element_text(hjust=0.5))
```


PLOT FOR SI APPENDIX (S15 Fig)

```{r}
# png('Exp2_TransferTestGazeEffects_All_Models.png', width=11.5, height=5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# grid.arrange(p1, p2, nrow=2, ncol=1)
# dev.off()
```


## Accuracy and mean RT as a function of EV differences, relative value differences, and gaze differences in the transfer test


1. Add the expected values and relative values for both options on each trial. Compute the EV difference and RV difference for the correct option minus the incorrect option. Also categorize proportional gaze for the correct option as < 33%, 33-67%, or > 67%.

```{r}
dat <- dat %>%
  mutate(left_EV = case_when(left_index==1 ~ 15,
                             left_index==2 ~ 18,
                             left_index==3 ~ 21,
                             left_index==4 ~ 24,
                             left_index==5 ~ 27,
                             left_index==6 ~ 30,
                             left_index==7 ~ 33,
                             left_index==8 ~ 36),
         right_EV = case_when(right_index==1 ~ 15,
                              right_index==2 ~ 18,
                              right_index==3 ~ 21,
                              right_index==4 ~ 24,
                              right_index==5 ~ 27,
                              right_index==6 ~ 30,
                              right_index==7 ~ 33,
                              right_index==8 ~ 36),
         left_RV = case_when(left_index==1 ~ 0,
                             left_index==2 ~ 1,
                             left_index==3 ~ 0,
                             left_index==4 ~ 1,
                             left_index==5 ~ 0,
                             left_index==6 ~ 1,
                             left_index==7 ~ 0,
                             left_index==8 ~ 1),
         right_RV = case_when(right_index==1 ~ 0,
                              right_index==2 ~ 1,
                              right_index==3 ~ 0,
                              right_index==4 ~ 1,
                              right_index==5 ~ 0,
                              right_index==6 ~ 1,
                              right_index==7 ~ 0,
                              right_index==8 ~ 1),
         EV_diff = ifelse(correct_resp == 'left', left_EV - right_EV, right_EV - left_EV),
         RV_diff = ifelse(correct_resp == 'left', left_RV - right_RV, right_RV - left_RV),
         prop_gaze_cor = case_when(gaze_cor < .33 ~ '<33%',
                                   gaze_cor >= .33 & gaze_cor <= .67 ~ '33-67%',
                                   gaze_cor > .67 ~ '>67%'))
         
```


2. Compute model-predicted accuracies on each trial for each subject. Merge with the data.


```{r}
# Step 1: For each subject, compute p(correct) on each trial across the 100 simulation runs
# Returns a 232 (trials) x 50 (subjects) matrix
model1_trialAcc <- sapply(sims1, function(X) rowMeans(X[,1,]), simplify='array')
model2_trialAcc <- sapply(sims2, function(X) rowMeans(X[,1,]), simplify='array')
model3_trialAcc <- sapply(sims3, function(X) rowMeans(X[,1,]), simplify='array')
model4_trialAcc <- sapply(sims4, function(X) rowMeans(X[,1,]), simplify='array')
model5_trialAcc <- sapply(sims5, function(X) rowMeans(X[,1,]), simplify='array')
model6_trialAcc <- sapply(sims6, function(X) rowMeans(X[,1,]), simplify='array')
model7_trialAcc <- sapply(sims7, function(X) rowMeans(X[,1,]), simplify='array')
model8_trialAcc <- sapply(sims8, function(X) rowMeans(X[,1,]), simplify='array')

# Step 2: Unwrap the model predicted accuracies and add them to the data
dat$model_1 <- c(model1_trialAcc)
dat$model_2 <- c(model2_trialAcc)
dat$model_3 <- c(model3_trialAcc)
dat$model_4 <- c(model4_trialAcc)
dat$model_5 <- c(model5_trialAcc)
dat$model_6 <- c(model6_trialAcc)
dat$model_7 <- c(model7_trialAcc)
dat$model_8 <- c(model8_trialAcc)
```


3. Average the data within each combination of EV difference, RV difference, and proportional gaze category.

```{r}
transferAccData <- dat %>%
  filter(block == 'transfer') %>% 
  group_by(EV_diff, RV_diff, prop_gaze_cor) %>% 
  summarise(prop_correct = mean(correct),
            model1_pred = mean(model_1),
            model2_pred = mean(model_2),
            model3_pred = mean(model_3),
            model4_pred = mean(model_4),
            model5_pred = mean(model_5),
            model6_pred = mean(model_6),
            model7_pred = mean(model_7),
            model8_pred = mean(model_8),
            n = n()) %>%
  ungroup() %>% 
  mutate(se = sqrt(prop_correct * (1 - prop_correct) / n),
         upper = prop_correct + se,
         lower = prop_correct - se)
```


4. Generate plot for accuracy (Fig 5I)

```{r}
# png('Exp2_TransferAccGazeEffects.png', width=5.5, height=2.7, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(data=transferAccData, aes(x=EV_diff, group=factor(RV_diff))) +
#   facet_wrap( ~ factor(prop_gaze_cor, levels=c('<33%', '33-67%', '>67%'),
#                        labels=c('Gaze correct < 33%', 'Gaze correct 33-67%', 'Gaze correct > 67%')),
#               nrow=1, ncol=3) +
#   geom_line(aes(y=prop_correct, col=factor(RV_diff))) +
#   geom_point(aes(y=prop_correct, col=factor(RV_diff))) +
#   geom_errorbar(aes(y=prop_correct, ymin=lower, ymax=upper, col=factor(RV_diff)), width=0.1) +
#   geom_line(aes(y=model8_pred, col=factor(RV_diff)), lty=2) +
#   geom_point(aes(y=model8_pred, col=factor(RV_diff)), pch=1) +
#   theme_classic(base_size=11) +
#   scale_y_continuous(limits=c(0,1)) +
#   scale_x_continuous(breaks=seq(0,21,3)) +
#   ylab('Choice accuracy') +
#   xlab('EV difference (correct - incorrect)') +
#   guides(col=guide_legend(title='Relative value diff.\n(correct - incorrect)')) +
#   scale_color_manual(values=c('maroon4', 'grey', 'maroon1')) +
#   theme(panel.spacing.x = unit(1, 'lines'),
#         legend.title.position = 'top',
#         legend.position = 'inside',
#         legend.position.inside = c(.85, .20),
#         legend.direction = 'horizontal',
#         legend.background = element_blank(),
#        # strip.background = element_blank(),
#         plot.title = element_text(size=11, hjust=0.5)) +
#   ggtitle('Exp. 2: Transfer')
# dev.off()
```


5. Compute model-predicted RTs on each trial for each subject. Merge with the data.


```{r}
# Step 1: For each subject, compute mean RT on each trial across the 100 simulation runs
# Returns a 232 (trials) x 50 (subjects) matrix
model1_trialRT <- sapply(sims1, function(X) rowMeans(X[,2,]), simplify='array')
model2_trialRT <- sapply(sims2, function(X) rowMeans(X[,2,]), simplify='array')
model3_trialRT <- sapply(sims3, function(X) rowMeans(X[,2,]), simplify='array')
model4_trialRT <- sapply(sims4, function(X) rowMeans(X[,2,]), simplify='array')
model5_trialRT <- sapply(sims5, function(X) rowMeans(X[,2,]), simplify='array')
model6_trialRT <- sapply(sims6, function(X) rowMeans(X[,2,]), simplify='array')
model7_trialRT <- sapply(sims7, function(X) rowMeans(X[,2,]), simplify='array')
model8_trialRT <- sapply(sims8, function(X) rowMeans(X[,2,]), simplify='array')

# Step 2: Unwrap the model predicted RTs and add them to the data
dat$model_1_RT <- c(model1_trialRT)
dat$model_2_RT <- c(model2_trialRT)
dat$model_3_RT <- c(model3_trialRT)
dat$model_4_RT <- c(model4_trialRT)
dat$model_5_RT <- c(model5_trialRT)
dat$model_6_RT <- c(model6_trialRT)
dat$model_7_RT <- c(model7_trialRT)
dat$model_8_RT <- c(model8_trialRT)
```


3. Average the data within each combination of EV difference, RV difference, and proportional gaze category.

```{r}
transferRTData <- dat %>%
  filter(RT >= 250, RT <= 10000) %>%
  filter(block == 'transfer') %>% 
  group_by(EV_diff, RV_diff, prop_gaze_cor) %>% 
  summarise(mean_RT = mean(RT),
            sd_RT = sd(RT),
            model1_pred = mean(model_1_RT),
            model2_pred = mean(model_2_RT),
            model3_pred = mean(model_3_RT),
            model4_pred = mean(model_4_RT),
            model5_pred = mean(model_5_RT),
            model6_pred = mean(model_6_RT),
            model7_pred = mean(model_7_RT),
            model8_pred = mean(model_8_RT),
            n = n()) %>%
  ungroup() %>% 
  mutate(se = sd_RT / sqrt(n),
         upper = mean_RT + se,
         lower = mean_RT - se)
```


4. Generate plot for mean RT (Fig 5J)

```{r}
# png('Exp2_TransferRTGazeEffects.png', width=5.5, height=2.7, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(data=transferRTData, aes(x=EV_diff, group=factor(RV_diff))) +
#   facet_wrap( ~ factor(prop_gaze_cor, levels=c('<33%', '33-67%', '>67%'),
#                        labels=c('Gaze correct < 33%', 'Gaze correct 33-67%', 'Gaze correct > 67%')),
#               nrow=1, ncol=3) +
#   geom_line(aes(y=mean_RT, col=factor(RV_diff)), show.legend=F) +
#   geom_point(aes(y=mean_RT, col=factor(RV_diff)), show.legend=F) +
#   geom_errorbar(aes(y=mean_RT, ymin=lower, ymax=upper, col=factor(RV_diff)), width=0.1, show.legend=F) +
#   geom_line(aes(y=model8_pred, col=factor(RV_diff)), lty=2, show.legend=F) +
#   geom_point(aes(y=model8_pred, col=factor(RV_diff)), pch=1, show.legend=F) +
#   theme_classic(base_size=11) +
#   #scale_y_continuous(limits=c(0,1)) +
#   scale_x_continuous(breaks=seq(0,21,3)) +
#   ylab('Mean RT (ms)') +
#   xlab('EV difference (correct - incorrect)') +
#   guides(col=guide_legend(title='Relative value diff.\n(correct - incorrect)')) +
#   scale_color_manual(values=c('maroon4', 'grey', 'maroon1')) +
#   theme(panel.spacing.x = unit(1, 'lines'),
#         legend.title.position = 'top',
#         legend.position = 'inside',
#         legend.position.inside = c(.85, .80),
#         legend.direction = 'horizontal',
#         legend.background = element_blank(),
#        # strip.background = element_blank(),
#         plot.title = element_text(size=11, hjust=0.5)) +
#   ggtitle('Exp. 2: Transfer')
# dev.off()
```



## Choice of left option as a function of relative gaze advantage and Q values

First add the model-predicted choice probabilities to the data

```{r}
# # Step 1: For each subject, compute p(correct) on each trial across the 100 simulation runs
# # Returns a 232 (trials) x 50 (subjects) matrix
# model1_trialAcc <- sapply(sims1, function(X) rowMeans(X[,1,]), simplify='array')
# model2_trialAcc <- sapply(sims2, function(X) rowMeans(X[,1,]), simplify='array')
# model3_trialAcc <- sapply(sims3, function(X) rowMeans(X[,1,]), simplify='array')
# model4_trialAcc <- sapply(sims4, function(X) rowMeans(X[,1,]), simplify='array')
# model5_trialAcc <- sapply(sims5, function(X) rowMeans(X[,1,]), simplify='array')
# model6_trialAcc <- sapply(sims6, function(X) rowMeans(X[,1,]), simplify='array')
# model7_trialAcc <- sapply(sims7, function(X) rowMeans(X[,1,]), simplify='array')
# model8_trialAcc <- sapply(sims8, function(X) rowMeans(X[,1,]), simplify='array')
# 
# # Step 2: Unwrap the model predicted accuracies and add them to the data
# dat$model_1 <- c(model1_trialAcc)
# dat$model_2 <- c(model2_trialAcc)
# dat$model_3 <- c(model3_trialAcc)
# dat$model_4 <- c(model4_trialAcc)
# dat$model_5 <- c(model5_trialAcc)
# dat$model_6 <- c(model6_trialAcc)
# dat$model_7 <- c(model7_trialAcc)
# dat$model_8 <- c(model8_trialAcc)
# 
# # Step 3: Convert from p(correct) to p(left)
# dat <- dat %>%
#   mutate(model_1 = ifelse(correct_resp == 'left', model_1, 1 - model_1),
#          model_2 = ifelse(correct_resp == 'left', model_2, 1 - model_2),
#          model_3 = ifelse(correct_resp == 'left', model_3, 1 - model_3),
#          model_4 = ifelse(correct_resp == 'left', model_4, 1 - model_4),
#          model_5 = ifelse(correct_resp == 'left', model_5, 1 - model_5),
#          model_6 = ifelse(correct_resp == 'left', model_6, 1 - model_6),
#          model_7 = ifelse(correct_resp == 'left', model_7, 1 - model_7),
#          model_8 = ifelse(correct_resp == 'left', model_8, 1 - model_8))
```


Next, add the latent Q values for the available options on each trial (Model 8)

```{r}
# sourceCpp('../../model_functions.cpp')
# source('../../model_utils.R')
# 
# # load fits for winning model (Model 8)
# model8_fits <- readRDS('../modeling/results/free-theta/model8_fits.RDS')
# 
# # extract fitted parameters
# model8_params <- sapply(model8_fits, function(X) X$optim$bestmem)
# 
# # latent variables from the model for each participant
# model8_latent <- lapply(seq_along(IDs), function(i) {
#   Simulate_Model(model8_params[,i], 
#                  model=8,
#                  options=modeling_data[[i]][['options']],
#                  outcomes=modeling_data[[i]][['outcomes']],
#                  avail=modeling_data[[i]][['avail']],
#                  gaze=modeling_data[[i]][['gaze']],
#                  n_opt=8,
#                  min_o=modeling_data[[i]][['min_o']],
#                  max_o=modeling_data[[i]][['max_o']],
#                  rt_lower=250,
#                  rt_upper=10000,
#                  delta=0,
#                  latent=TRUE)
})
```


```{r}
# dat$Q1 <- NA
# dat$Q2 <- NA
# 
# for ( i in seq_along(IDs) ) {
#   dat$Q1[dat$subject == IDs[i]] <- model8_latent[[i]][[4]][,1]
#   dat$Q2[dat$subject == IDs[i]] <- model8_latent[[i]][[4]][,2]
# }
```


For each subject, divide the learning and transfer trials into 5 quintiles based on Q-value differences and gaze differences (left-right)

```{r}
# dat <- dat %>%
#   mutate(Q_diff = Q1 - Q2,
#          gaze_diff_LR = pre_fix_left - pre_fix_right) %>%
#   group_by(subject, block) %>%
#   mutate(Q_diff_bucket = ntile(Q_diff, 5),
#          gaze_diff_LR_bucket = ntile(gaze_diff_LR, 5)) %>%
#   ungroup()
```


Compute proportion of left choices within each combination of Q-gaze quintiles. Then average across participants.

```{r}
# df1 <- dat %>%
#   group_by(subject, block, Q_diff_bucket, gaze_diff_LR_bucket) %>%
#   summarise(n = n(),
#             prop_left = mean(choice_id == 1),
#             model = mean(model_8),
#             min_Q_diff = min(Q_diff),
#             max_Q_diff = max(Q_diff)) %>%
#   ungroup()
# 
# df2 <- df1 %>%
#   group_by(block, Q_diff_bucket, gaze_diff_LR_bucket) %>%
#   summarise(mean_prop_left = mean(prop_left),
#             sd_prop_left = sd(prop_left),
#             n = n(),
#             avg_min_Q_diff = mean(min_Q_diff),
#             avg_max_Q_diff = mean(max_Q_diff),
#             model = mean(model)) %>%
#   ungroup() %>%
#   mutate(se = sd_prop_left / sqrt(n),
#          lower = mean_prop_left - se,
#          upper = mean_prop_left + se,
#          Q_diff_bucket = factor(Q_diff_bucket),
#          gaze_diff_LR_bucket = factor(gaze_diff_LR_bucket),
#          block = factor(block, levels=c('learning', 'transfer'),
#                         labels=c('Exp. 2: Learning', 'Exp. 2: Transfer')))
```



```{r}
# p1 <- ggplot(df2, aes(x=gaze_diff_LR_bucket, y=mean_prop_left, group=Q_diff_bucket)) +
#   facet_wrap(~ block, nrow=1, ncol=2) +
#   geom_hline(yintercept=0.5, lty=3) +
#   geom_line(aes(col=Q_diff_bucket)) +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, col='grey40') +
#   geom_point(aes(fill=Q_diff_bucket), pch=21, col='black') +
#   geom_line(aes(y=model, col=Q_diff_bucket), lty=2, show.legend=F) +
#   theme_bw() +
#   theme(panel.grid = element_blank(),
#         strip.background = element_blank(),
#         plot.title = element_text(size=24),
#         legend.text = element_text(size=8),
#         legend.title = element_text(size=8),
#         legend.key.size = unit(0.1, 'cm'),
#         legend.key.spacing.y = unit(0.05, 'cm')) +
#   scale_color_viridis_d(option='viridis') +
#   scale_fill_viridis_d(option='viridis') +
#   labs(x='Left gaze \u2212 Right gaze (quintiles)', y='Proportion of Left Choices') +
#   guides(col=guide_legend(title='Left Q \u2212 Right Q\n(quintiles)'), 
#          fill=guide_legend(title='Left Q \u2212 Right Q\n(quintiles)')) +
#   ggtitle('A')
```


Filter out all trials in which |Q_left - Q_right| >= 0.1. Then divide the remaining trials for each subject into five quintiles based on average Q-value magnitude.

Compute the proportion of left choices within each combination of magnitude-gaze quintiles. Then average across participants.

```{r}
# df3 <- dat %>%
#   filter(abs(Q_diff) < 0.1) %>%
#   mutate(Q_mag = (Q1 + Q2) / 2) %>%
#   group_by(subject) %>%
#   mutate(Q_mag_bucket = ntile(Q_mag, 5),
#          gaze_diff_LR_bucket = ntile(gaze_diff_LR, 5)) %>%
#   ungroup()
# 
# df4 <- df3 %>%
#   group_by(subject, Q_mag_bucket, gaze_diff_LR_bucket) %>%
#   summarise(n = n(),
#             prop_left = mean(choice_id == 1),
#             model = mean(model_8),
#             min_Q_diff = min(Q_diff),
#             max_Q_diff = max(Q_diff)) %>%
#   ungroup()
# 
# df5 <- df4 %>%
#   group_by(Q_mag_bucket, gaze_diff_LR_bucket) %>%
#   summarise(mean_prop_left = mean(prop_left),
#             sd_prop_left = sd(prop_left),
#             n = n(),
#             avg_min_Q_diff = mean(min_Q_diff),
#             avg_max_Q_diff = mean(max_Q_diff),
#             model = mean(model)) %>%
#   ungroup() %>%
#   mutate(se = sd_prop_left / sqrt(n),
#          lower = mean_prop_left - se,
#          upper = mean_prop_left + se,
#          Q_mag_bucket = factor(Q_mag_bucket),
#          gaze_diff_LR_bucket = factor(gaze_diff_LR_bucket))
```


```{r}
# p2 <- ggplot(df5, aes(x=gaze_diff_LR_bucket, y=mean_prop_left, group=Q_mag_bucket)) +
#   geom_hline(yintercept=0.5, lty=3) +
#   geom_line(aes(col=Q_mag_bucket)) +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.1, col='grey40') +
#   geom_point(aes(fill=Q_mag_bucket), pch=21, col='black') +
#   geom_line(aes(y=model, col=Q_mag_bucket), lty=2, show.legend=F) +
#   theme_bw() +
#   theme(panel.grid = element_blank(),
#         plot.title = element_text(size=24),
#         legend.position = 'inside',
#         legend.position.inside = c(.8, .25),
#         legend.text = element_text(size=8),
#         legend.title = element_text(size=8),
#         legend.key.size = unit(0.1, 'cm'),
#         legend.key.spacing.y = unit(0.05, 'cm')) +
#   scale_color_viridis_d(option='viridis') +
#   scale_fill_viridis_d(option='viridis') +
#   labs(x='Left gaze \u2212 Right gaze (quintiles)', y='Proportion of Left Choices') +
#   guides(col=guide_legend(title='Q magnitude\n(quintiles)', byrow=T), 
#          fill=guide_legend(title='Q magnitude\n(quintiles)', byrow=T)) +
#   scale_y_continuous(limits=c(0,1)) +
#   ggtitle('B')
```


```{r}
# tiff('Exp2_gaze_Qvalues.tiff', width=8.5, height=3.5, units='in', res=300)
# grid.arrange(p1, p2, layout_matrix=rbind(c(1,1,2)))
# dev.off()
```



## Model comparison 

First, we need to add the models with theta as a free parameter.

```{r}
model_order <- c('model_1', 'model_2', 'model_3',
                 'model_4', 'model_4FT', 'model_5', 'model_5FT', 
                 'model_6', 'model_6FT', 'model_7', 'model_7FT',
                 'model_8', 'model_8FT')

model_labels <- c('Q',
                  'Q * gaze',
                  'Q + gaze',
                  'softmax(Q)',
                  'softmax(Q)',
                  'softmax(Q * gaze)',
                  'softmax(Q * gaze)',
                  'softmax(Q) * gaze',
                  'softmax(Q) * gaze',
                  'softmax(Q) + gaze',
                  'softmax(Q) + gaze',
                  'softmax(Q + gaze)',
                  'softmax(Q + gaze)')

model_version <- c('no theta', 'no theta', 'no theta',
                   'theta = 50', 'theta free', 'theta = 50', 'theta free',
                   'theta = 50', 'theta free', 'theta = 50', 'theta free',
                   'theta = 50', 'theta free')
```

1. Load accumulative one-step-ahead prediction error (APE) results for each model. These were run in batches.

```{r}
model1_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/model1_batch', i, '_APE.RDS'))
})

model2_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/model2_batch', i, '_APE.RDS'))
})

model3_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/model3_batch', i, '_APE.RDS'))
})

model4_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/model4_batch', i, '_APE.RDS'))
})

model5_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/model5_batch', i, '_APE.RDS'))
})

model6_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/model6_batch', i, '_APE.RDS'))
})

model7_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/model7_batch', i, '_APE.RDS'))
})

model8_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/model8_batch', i, '_APE.RDS'))
})

model4FT_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/free-theta/model4_batch', i, '_APE.RDS'))
})

model5FT_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/free-theta/model5_batch', i, '_APE.RDS'))
})

model6FT_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/free-theta/model6_batch', i, '_APE.RDS'))
})

model7FT_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/free-theta/model7_batch', i, '_APE.RDS'))
})

model8FT_APE_results <- lapply(1:20, FUN=function(i) {
  readRDS(paste0('../modeling/results/free-theta/model8_batch', i, '_APE.RDS'))
})


model1_APE_results <- unlist(model1_APE_results, recursive = FALSE)
model2_APE_results <- unlist(model2_APE_results, recursive = FALSE)
model3_APE_results <- unlist(model3_APE_results, recursive = FALSE)
model4_APE_results <- unlist(model4_APE_results, recursive = FALSE)
model5_APE_results <- unlist(model5_APE_results, recursive = FALSE)
model6_APE_results <- unlist(model6_APE_results, recursive = FALSE)
model7_APE_results <- unlist(model7_APE_results, recursive = FALSE)
model8_APE_results <- unlist(model8_APE_results, recursive = FALSE)
model4FT_APE_results <- unlist(model4FT_APE_results, recursive = FALSE)
model5FT_APE_results <- unlist(model5FT_APE_results, recursive = FALSE)
model6FT_APE_results <- unlist(model6FT_APE_results, recursive = FALSE)
model7FT_APE_results <- unlist(model7FT_APE_results, recursive = FALSE)
model8FT_APE_results <- unlist(model8FT_APE_results, recursive = FALSE)
```


2. Extract the one-step-ahead log-likelihoods, recode infinite LLs as NAs (there were very few of these), and compute the difference in APE relative to the winning model (softmax(Q) + gaze) for each participant.

```{r}
# extract one-step-ahead LLs (trials x subjects matrices)
model1_LL <- sapply(model1_APE_results, function(X) X$LL)
model2_LL <- sapply(model2_APE_results, function(X) X$LL)
model3_LL <- sapply(model3_APE_results, function(X) X$LL)
model4_LL <- sapply(model4_APE_results, function(X) X$LL)
model5_LL <- sapply(model5_APE_results, function(X) X$LL)
model6_LL <- sapply(model6_APE_results, function(X) X$LL)
model7_LL <- sapply(model7_APE_results, function(X) X$LL)
model8_LL <- sapply(model8_APE_results, function(X) X$LL)
model4FT_LL <- sapply(model4FT_APE_results, function(X) X$LL)
model5FT_LL <- sapply(model5FT_APE_results, function(X) X$LL)
model6FT_LL <- sapply(model6FT_APE_results, function(X) X$LL)
model7FT_LL <- sapply(model7FT_APE_results, function(X) X$LL)
model8FT_LL <- sapply(model8FT_APE_results, function(X) X$LL)

# recode infinite LLs
model1_LL[is.infinite(model1_LL)] <- NA
model2_LL[is.infinite(model2_LL)] <- NA
model3_LL[is.infinite(model3_LL)] <- NA
model4_LL[is.infinite(model4_LL)] <- NA
model5_LL[is.infinite(model5_LL)] <- NA
model6_LL[is.infinite(model6_LL)] <- NA
model7_LL[is.infinite(model7_LL)] <- NA
model8_LL[is.infinite(model8_LL)] <- NA
model4FT_LL[is.infinite(model4FT_LL)] <- NA
model5FT_LL[is.infinite(model5FT_LL)] <- NA
model6FT_LL[is.infinite(model6FT_LL)] <- NA
model7FT_LL[is.infinite(model7FT_LL)] <- NA
model8FT_LL[is.infinite(model8FT_LL)] <- NA

# trim off first five trials in Models 1-7 (all zeros since APE was computed based on trials 6-232)
# the code for computing APE was updated before running the other models.
model1_LL <- model1_LL[-c(1:5),]
model2_LL <- model2_LL[-c(1:5),]
model3_LL <- model3_LL[-c(1:5),]
model4_LL <- model4_LL[-c(1:5),]
model5_LL <- model5_LL[-c(1:5),]
model6_LL <- model6_LL[-c(1:5),]
model7_LL <- model7_LL[-c(1:5),]

# compute difference in APE relative to Model 8 w/ theta parameter free
# doing it this way ensures that if one model produced an infinite LL on a particular trial, 
# the trial will not be counted for either of the models being compared (both will have NA value for that trial)  
N <- length(IDs)
APE_data <- data.frame(model = rep(model_order, each=N),
                       version = rep(model_version, each=N),
                       subject = rep(IDs, length(model_order)),
                       APE_diff = c( colSums( (-model1_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model2_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model3_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model4_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model4FT_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model5_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model5FT_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model6_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model6FT_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model7_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model7FT_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model8_LL) - (-model8FT_LL) , na.rm=TRUE),
                                     colSums( (-model8FT_LL) - (-model8FT_LL) , na.rm=TRUE)))

APE_diff_data <- APE_data %>%
  mutate(model = factor(model, levels=model_order, labels=model_labels),
         version = factor(version, levels=c('no theta', 'theta = 50', 'theta free'))) %>%
  group_by(model, version) %>%
  summarise(mean_APE_diff = mean(APE_diff),
            sd_APE_diff = sd(APE_diff),
            n = n()) %>%
  ungroup() %>%
  mutate(se = sd_APE_diff / sqrt(n),
         lower = mean_APE_diff - se,
         upper = mean_APE_diff + se,
         gaze_model = factor(ifelse(model %in% c('Q','softmax(Q)'), 0, 1)))

APE_diff_data$label <- c('***', '***', '***', '***', '***', '***', '***', '***', '***', '***', '**', '***', '')

```

PLOT FOR MANUSCRIPT (Fig 2A)

```{r}
# png('APE_Exp2.png', width=3.5, height=3, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(subset(APE_diff_data, version %in% c('no theta', 'theta free')), aes(x=model, y=mean_APE_diff)) +
#   geom_bar(fill='grey40', col='black', stat='identity', width=0.5) +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
#   ylab(expression(paste(Delta, 'APE (model i ', '\U2212', ' model 8)'))) +
#   xlab('Model') +
#   theme_classic(base_size=11) +
#   theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   scale_y_continuous(limits=c(0, 60), expand=c(0,0)) +
#   geom_text(aes(y=upper+5, label=label)) +
#   ggtitle('Exp. 2')
# dev.off()
```

```{r}
with(APE_data, t.test(APE_diff[model=='model_1']))
```

```{r}
with(APE_data, t.test(APE_diff[model=='model_2']))
```

```{r}
with(APE_data, t.test(APE_diff[model=='model_3']))
```

```{r}
with(APE_data, t.test(APE_diff[model=='model_4']))
```

```{r}
with(APE_data, t.test(APE_diff[model=='model_4FT']))
```

```{r}
with(APE_data, t.test(APE_diff[model=='model_5']))
```
```{r}
with(APE_data, t.test(APE_diff[model=='model_5FT']))
```

```{r}
with(APE_data, t.test(APE_diff[model=='model_6']))
```
```{r}
with(APE_data, t.test(APE_diff[model=='model_6FT']))
```
```{r}
with(APE_data, t.test(APE_diff[model=='model_7']))
```


```{r}
with(APE_data, t.test(APE_diff[model=='model_7FT']))
```
```{r}
with(APE_data, t.test(APE_diff[model=='model_8']))
```


PLOT WITH EXTENDED SET OF MODELS

```{r}
# png('APE_Exp2_extended.png', width=6.5, height=3.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(APE_diff_data, aes(x=model, y=mean_APE_diff, fill=version)) +
#   geom_hline(yintercept = 0, lty=2) +
#   geom_col(position=position_dodge(width=0.6), width=0.5, col='black') +
#   geom_errorbar(position=position_dodge(width=0.6), aes(ymin=lower, ymax=upper), width=0.2) +
#   ylab(expression(paste(Delta, 'APE (model i ', '\U2212', ' model 8 [theta free])'))) +
#   xlab('Model') +
#   theme_classic(base_size=11) +
#   theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1),
#         plot.title = element_text(hjust=0.5, size=11)) +
#   scale_y_continuous(limits=c(0, 60), breaks=seq(0, 60, 10), expand=c(0,0)) +
#   geom_text(aes(y=upper+1.5, label=label), position=position_dodge(width=0.6)) +
#   ggtitle('Exp. 2') +
#   scale_fill_manual(values=c('no theta' = 'white',
#                              'theta = 50' = 'grey80',
#                              'theta free' = 'grey40')) +
#   guides(fill = guide_legend(title=''))
# dev.off()
```



## Figure 2B: Model accuracy

1. Compute model-predicted accuracy on each trial for each subject. Merge with the data.

```{r}
# Step 1: For each subject, compute p(correct) on each trial across the 100 simulation runs
# Returns a 232 (trials) x 50 (subjects) matrix
model1_trialAcc <- sapply(sims1, function(X) rowMeans(X[,1,]), simplify='array')
model2_trialAcc <- sapply(sims2, function(X) rowMeans(X[,1,]), simplify='array')
model3_trialAcc <- sapply(sims3, function(X) rowMeans(X[,1,]), simplify='array')
model4_trialAcc <- sapply(sims4, function(X) rowMeans(X[,1,]), simplify='array')
model5_trialAcc <- sapply(sims5, function(X) rowMeans(X[,1,]), simplify='array')
model6_trialAcc <- sapply(sims6, function(X) rowMeans(X[,1,]), simplify='array')
model7_trialAcc <- sapply(sims7, function(X) rowMeans(X[,1,]), simplify='array')
model8_trialAcc <- sapply(sims8, function(X) rowMeans(X[,1,]), simplify='array')

# Step 2: Unwrap the model predicted accuracies and add them to the data
dat$model_1 <- c(model1_trialAcc)
dat$model_2 <- c(model2_trialAcc)
dat$model_3 <- c(model3_trialAcc)
dat$model_4 <- c(model4_trialAcc)
dat$model_5 <- c(model5_trialAcc)
dat$model_6 <- c(model6_trialAcc)
dat$model_7 <- c(model7_trialAcc)
dat$model_8 <- c(model8_trialAcc)
```



2. Classify model predictions as correct or incorrect

```{r}
dat$model1_correct <- ifelse((dat$correct == 1 & dat$model_1 > 0.5) | (dat$correct == 0 & dat$model_1 < 0.5), 1, 0)
dat$model2_correct <- ifelse((dat$correct == 1 & dat$model_2 > 0.5) | (dat$correct == 0 & dat$model_2 < 0.5), 1, 0)
dat$model3_correct <- ifelse((dat$correct == 1 & dat$model_3 > 0.5) | (dat$correct == 0 & dat$model_3 < 0.5), 1, 0)
dat$model4_correct <- ifelse((dat$correct == 1 & dat$model_4 > 0.5) | (dat$correct == 0 & dat$model_4 < 0.5), 1, 0)
dat$model5_correct <- ifelse((dat$correct == 1 & dat$model_5 > 0.5) | (dat$correct == 0 & dat$model_5 < 0.5), 1, 0)
dat$model6_correct <- ifelse((dat$correct == 1 & dat$model_6 > 0.5) | (dat$correct == 0 & dat$model_6 < 0.5), 1, 0)
dat$model7_correct <- ifelse((dat$correct == 1 & dat$model_7 > 0.5) | (dat$correct == 0 & dat$model_7 < 0.5), 1, 0)
dat$model8_correct <- ifelse((dat$correct == 1 & dat$model_8 > 0.5) | (dat$correct == 0 & dat$model_8 < 0.5), 1, 0)
```


3. Averge model prediction accuracy (all trials)

```{r}
modelPredAccData <- dat %>%
  group_by(subject) %>%
  summarise(model_1 = mean(model1_correct),
            model_2 = mean(model2_correct),
            model_3 = mean(model3_correct),
            model_4 = mean(model4_correct),
            model_5 = mean(model5_correct),
            model_6 = mean(model6_correct),
            model_7 = mean(model7_correct),
            model_8 = mean(model8_correct),
            n = n()) %>%
  ungroup()

modelPredAccData1 <- modelPredAccData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='accuracy')

# order of models
model_order <- c('model_1', 'model_2', 'model_3', 'model_4',
                 'model_5', 'model_6', 'model_7', 'model_8')

modelPredAccData2 <- modelPredAccData1 %>%
  group_by(model) %>%
  summarise(mean_accuracy = mean(accuracy),
            sd_accuracy = sd(accuracy),
            n = n()) %>%
  ungroup() %>%
  mutate(se_acc = sd_accuracy / sqrt(n),
         lower = mean_accuracy - se_acc,
         upper = mean_accuracy + se_acc,
         model = factor(model, levels=model_order, labels=c('1','2','3','4','5','6','7','8*')),
         gaze_model = factor(ifelse(model %in% c('1','4'), 0, 1)))
```


4. PLOT FOR MANUSCRIPT (MODEL PREDICTION ACCURACY; Fig 2B)

```{r}
# png('Exp2_model_prediction_accuracy.png', width=2, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(modelPredAccData2, aes(x=model, y=mean_accuracy)) +
#   geom_bar(aes(fill=gaze_model), col='black', stat='identity', width=0.6, show.legend=FALSE) +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
#   ylab('Model prediction acc.') +
#   xlab('Model') +
#   theme_classic(base_size=11) +
#   theme(plot.title = element_text(hjust=0.5, size=11)) +
#   scale_y_continuous(limits=c(0, 1), expand=c(0,0)) +
#   ggtitle('Exp. 2') +
#   scale_fill_manual(values=c('grey80', 'grey40'))
# dev.off()
```



5. Average model prediction accuracy within trial categories:

* Correct choice + looked longer at correct option
* Correct choice + looked longer at incorrect option
* Incorrect choice + looked longer at correct option
* Incorrect choice + looked longer at incorrect option

```{r}
modelPredAccData <- dat %>%
  mutate(correct_gaze_adv = ifelse(gaze_cor > gaze_inc, 1, 0)) %>%
  group_by(subject, correct, correct_gaze_adv) %>%
  summarise(model_1 = mean(model1_correct),
            model_2 = mean(model2_correct),
            model_3 = mean(model3_correct),
            model_4 = mean(model4_correct),
            model_5 = mean(model5_correct),
            model_6 = mean(model6_correct),
            model_7 = mean(model7_correct),
            model_8 = mean(model8_correct),
            n = n()) %>%
  ungroup() %>%
  group_by(subject) %>%
  mutate(total = sum(n),
         prop = n / total) %>%
  ungroup()

modelPredAccData1 <- modelPredAccData %>%
  pivot_longer(cols=model_1:model_8,
               names_to='model',
               values_to='accuracy')

# order of models
model_order <- c('model_1', 'model_2', 'model_3', 'model_4',
                 'model_5', 'model_6', 'model_7', 'model_8')

modelPredAccData2 <- modelPredAccData1 %>%
  group_by(correct, correct_gaze_adv, model) %>%
  summarise(mean_accuracy = mean(accuracy),
            sd_accuracy = sd(accuracy),
            mean_prop = mean(prop),
            sd_prop = sd(prop),
            n = n()) %>%
  ungroup() %>%
  mutate(se_acc = sd_accuracy / sqrt(n),
         lower = mean_accuracy - se_acc,
         upper = mean_accuracy + se_acc,
         se_prop = sd_prop / sqrt(n),
         model = factor(model, levels=model_order, labels=c('1','2','3','4','5','6','7','8*')),
         correct = factor(correct, levels=c(0,1), labels=c('Incorrect choice', 'Correct choice')),
         correct_gaze_adv = factor(correct_gaze_adv, levels=c(0,1), 
                                   labels=c('Gaze: Inc. > Cor.', 'Gaze: Inc. < Cor.')),
         gaze_model = factor(ifelse(model %in% c('1','4'), 0, 1)))
```


6. PLOT FOR MANUSCRIPT (MODEL PREDICTION ACCURACY BY TRIAL TYPE; Fig 2C)

```{r}
# png('Exp2_model_prediction_accuracy_by_gaze.png', width=6, height=2.5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# ggplot(modelPredAccData2, aes(x=model, y=mean_accuracy)) +
#   facet_wrap(~ correct + correct_gaze_adv, nrow=1, ncol=4) +
#   geom_bar(aes(fill=gaze_model), col='black', stat='identity', width=0.6, show.legend=FALSE) +
#   geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2) +
#   ylab('Model prediction acc.') +
#   xlab('Model') +
#   theme_classic(base_size=11) +
#   theme(plot.title = element_text(hjust=0.5, size=11),
#         strip.background = element_blank()) +
#   scale_y_continuous(limits=c(0, 1), expand=c(0,0)) +
#   ggtitle('Exp. 2') +
#   scale_fill_manual(values=c('grey80', 'grey40'))
# dev.off()
```






## RL-SSM latent variables for an example participant

1. Load model functions and estimated parameters. Then, extract latent Q values and mean drift rates for example participant.

```{r}
sourceCpp('../../model_functions.cpp')

# load fits
model8_fits <- readRDS('../modeling/results/free-theta/model8_fits.RDS')

# extract fitted parameters
model8_params <- sapply(model8_fits, function(X) X$optim$bestmem)

# parameters for example participant
subject32_model8_params <- model8_params[, IDs == 32]
#subject43_model8_params <- model8_params[, IDs == 43]

# extract latent variables
subject32_model8_latentVars <- Simulate_Model(subject32_model8_params, 
                                              model=8,
                                              options=modeling_data[[which(IDs == 32)]][['options']],
                                              outcomes=modeling_data[[which(IDs == 32)]][['outcomes']],
                                              avail=modeling_data[[which(IDs == 32)]][['avail']],
                                              gaze=modeling_data[[which(IDs == 32)]][['gaze']],
                                              n_opt=8,
                                              min_o=modeling_data[[which(IDs == 32)]][['min_o']],
                                              max_o=modeling_data[[which(IDs == 32)]][['max_o']],
                                              rt_lower=250,
                                              rt_upper=10000,
                                              delta=0,
                                              latent=TRUE)
```

2. Put latent variables into data frames.

```{r}
which_correct <- datasets[[which(IDs == 32)]][['correct_resp']] # which option was the correct choice on each trial
correct_choice <- datasets[[which(IDs == 32)]][['correct']] # whether the participant chose the correct option on each trial

# Q values for available options
Q_value_df <- data.frame(model='model_8',
                         option=rep(c('correct','incorrect'), each=120),
                         trial=rep(1:120, times=2),
                         correct=rep(correct_choice[1:120], times=2),
                         Q=c(subject32_model8_latentVars[[4]][cbind(1:120, ifelse(which_correct[1:120]=='left', 1, 2))],
                             subject32_model8_latentVars[[4]][cbind(1:120, ifelse(which_correct[1:120]=='left', 2, 1))]))

# mean drift rates
dat$model8_correct <- ifelse((dat$correct == 1 & dat$model_8 > 0.5) | (dat$correct == 0 & dat$model_8 <= 0.5), 1, 0)
drift_df <- data.frame(model='model_8',
                       option=rep(c('correct','incorrect'), each=120),
                       trial=rep(1:120, times=2),
                       correct=rep(correct_choice[1:120], times=2),
                       driftMean=c(subject32_model8_latentVars[[5]][cbind(1:120, ifelse(which_correct[1:120]=='left', 1, 2))],
                                   subject32_model8_latentVars[[5]][cbind(1:120, ifelse(which_correct[1:120]=='left', 2, 1))]),
                       model_correct=c(with(subset(dat, subject==32), model8_correct[1:120]),
                                       with(subset(dat, subject==32), model8_correct[1:120])))

drift_df$label <- ifelse(drift_df$model_correct == 1, '', '0')

# proportional gaze
gaze_df <- data.frame(model='model_8',
                      option=rep(c('correct','incorrect'), each=120),
                      trial=rep(1:120, times=2),
                      gaze=c(modeling_data[[which(IDs == 32)]][['gaze']][cbind(1:120, ifelse(which_correct[1:120]=='left', 1, 2))],
                             modeling_data[[which(IDs == 32)]][['gaze']][cbind(1:120, ifelse(which_correct[1:120]=='left', 2, 1))]))
```


3. Generate plot (S3 Fig)

```{r}
rect_data <- data.frame(trial=1:120, correct=correct_choice[1:120]) %>%
  mutate(
    xmin = trial - 0.5,
    xmax = trial + 0.5,
    fill_color = ifelse(correct == 1, "green", "red")
  )


p1 <- ggplot(Q_value_df, aes(x = trial, y = Q, group = option)) +
  # Add shaded backgrounds
  geom_rect(data = rect_data,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill_color),
            inherit.aes = FALSE,
            alpha = 0.2) +
  geom_point(aes(pch=option), show.legend=FALSE, cex=0.5) +
  scale_shape_manual(values=c(16, 4)) +
  scale_fill_identity() +  # Use colors directly from fill_color
  theme_minimal(base_size=11) +
  theme(panel.grid = element_blank()) +
  labs(x = "", y = "Q value") +
  #scale_y_continuous(limits=c(.42, .58), breaks=c(.42, .46, .50, .54, .58)) +
  scale_x_continuous(limits=c(0, 121), breaks=c(1, 15, 30, 45, 60, 75, 90, 105, 120), expand=c(0,0))


p2 <- ggplot(gaze_df, aes(x = trial, y = gaze, group = option)) +
  # Add shaded backgrounds
  geom_rect(data = rect_data,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill_color),
            inherit.aes = FALSE,
            alpha = 0.2) +
  geom_point(aes(pch=option), show.legend=FALSE, cex=0.5) +
  scale_shape_manual(values=c(16, 4)) +
  scale_fill_identity() +  # Use colors directly from fill_color
  theme_minimal(base_size=11) +
  theme(panel.grid = element_blank()) +
  labs(x = "", y = "Prop. gaze") +
  scale_y_continuous(limits=c(0, 1)) + #breaks=c(.42, .46, .50, .54, .58)) +
  scale_x_continuous(limits=c(0, 121), breaks=c(1, 15, 30, 45, 60, 75, 90, 105, 120), expand=c(0,0))



p3 <- ggplot(drift_df, aes(x = trial, y = driftMean, group = option)) +
  # Add shaded backgrounds
  geom_rect(data = rect_data,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill_color),
            inherit.aes = FALSE,
            alpha = 0.2) +
  geom_point(aes(pch=option), show.legend=FALSE, cex=0.5) +
  scale_shape_manual(values=c(16, 4)) +
  scale_fill_identity() +  # Use colors directly from fill_color
  theme_minimal(base_size=11) +
  theme(panel.grid = element_blank()) +
  labs(x = "Trial number", y = "Mean drift rate") +
  scale_y_continuous(limits=c(0, .40), labels=function(x) sprintf("%.2f", x)) +
  scale_x_continuous(limits=c(0, 121), breaks=c(1, 15, 30, 45, 60, 75, 90, 105, 120), expand=c(0,0)) +
  geom_text(aes(y=.40, label=label), size=2, col='red')
```



```{r}
# png('Exp2_Subject32.png', width=6.5, height=5, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# grid.arrange(p1, p2, p3, nrow=3, ncol=1)
# dev.off()
```




## Figure 3A: Distributions of estimated parameters for the softmax(Q) + gaze model

```{r}
Exp1_fits <- readRDS('../../Exp1_TwoContexts/modeling/results/free-theta/model7_fits.RDS')
Exp2_fits <- readRDS('../modeling/results/free-theta/model7_fits.RDS')

Exp1_params <- sapply(Exp1_fits, function(X) X$optim$bestmem)
Exp2_params <- sapply(Exp2_fits, function(X) X$optim$bestmem)


#png('parameter_estimates_Model7.png', width=14, height=2, units='in', res=300)
par(mfrow=c(1,8), mar=c(4.5,1,3,1))
plot(density(Exp1_params[2,]), col='maroon1', xaxs='i', yaxs='i',  xlim=c(0, 1), ylim=c(0, 25), xlab=expression(alpha), ylab='', yaxt='n',
     main='Learning\nrate', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[2,]), col='maroon4', lty=2)
mtext(paste0('Exp. 1: M = ', format(round(mean(Exp1_params[2,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('Exp. 2: M = ', format(round(mean(Exp2_params[2,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp2_params[7,]), col='maroon4', lty=2, xaxs='i',  yaxs='i', xlim=c(0, 1), ylim=c(0, 1.5), xlab=expression(italic(w)['rel']), ylab='', yaxt='n', main='Relative\nencoding', cex.lab=1.8, cex.main=1.5, las=2)
mtext(paste0('M = ', format(round(mean(Exp2_params[7,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[3,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 1.2), ylim=c(0, 4.7), xlab=expression(italic(beta)['Q']), ylab='', yaxt='n', main='Q drift\nscaling', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[3,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[3,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[3,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[4,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 1.2), ylim=c(0, 4.7), xlab=expression(italic(beta)['gaze']), ylab='', yaxt='n', main='Gaze drift\nscaling', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[4,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[4,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[4,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[7,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 80), #ylim=c(0, 4.7), 
     xlab=expression(italic(theta)), ylab='', yaxt='n', main='Softmax\ninv. temp.', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[8,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[7,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[8,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[6,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 1500), ylim=c(0, .0035), xlab=expression(italic(A)), ylab='', yaxt='n',
     main='Start point\nupper bound', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[6,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[6,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[6,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[6,] + Exp1_params[5,]), col='maroon1', xaxs='i', yaxs='i', ylim=c(0, .003), xlab=expression(italic(b)), ylab='', yaxt='n',
     main='Decision\nthreshold', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[6,] + Exp2_params[5,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[6,] + Exp1_params[5,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[6,] + Exp2_params[5,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[1,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 300), ylim=c(0, .06), xlab=expression(italic(t)[0]), ylab='', yaxt='n',
     main='Non-decision\ntime', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[1,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[1,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.55, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[1,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.55, col='maroon4', cex=0.75)
#dev.off()
```


## Figure 3B: Distributions of estimated parameters for the softmax(Q + gaze) model

```{r}
Exp1_fits <- readRDS('../../Exp1_TwoContexts/modeling/results/free-theta/model8_fits.RDS')
Exp2_fits <- readRDS('../modeling/results/free-theta/model8_fits.RDS')

Exp1_params <- sapply(Exp1_fits, function(X) X$optim$bestmem)
Exp2_params <- sapply(Exp2_fits, function(X) X$optim$bestmem)


#png('parameter_estimates_Model8.png', width=14, height=2, units='in', res=300)
par(mfrow=c(1,8), mar=c(4.5,1,3,1))
plot(density(Exp1_params[2,]), col='maroon1', xaxs='i', yaxs='i',  xlim=c(0, 1), ylim=c(0, 25), xlab=expression(alpha), ylab='', yaxt='n',
     main='Learning\nrate', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[2,]), col='maroon4', lty=2)
mtext(paste0('Exp. 1: M = ', format(round(mean(Exp1_params[2,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('Exp. 2: M = ', format(round(mean(Exp2_params[2,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp2_params[7,]), col='maroon4', lty=2, xaxs='i',  yaxs='i', xlim=c(0, 1), ylim=c(0, 1.5), xlab=expression(italic(w)['rel']), ylab='', yaxt='n', main='Relative\nencoding', cex.lab=1.8, cex.main=1.5, las=2)
mtext(paste0('M = ', format(round(mean(Exp2_params[7,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[3,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 1.2), ylim=c(0, 10), xlab=expression(italic(beta)['Q']), ylab='', yaxt='n', main='Q drift\nscaling', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[3,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[3,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[3,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[4,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 1.2), ylim=c(0, 5), xlab=expression(italic(beta)['gaze']), ylab='', yaxt='n', main='Gaze drift\nscaling', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[4,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[4,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[4,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[7,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 80), ylim=c(0, .2), 
     xlab=expression(italic(theta)), ylab='', yaxt='n', main='Softmax\ninv. temp.', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[8,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[7,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[8,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[6,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 1500), ylim=c(0, .0045), xlab=expression(italic(A)), ylab='', yaxt='n',
     main='Start point\nupper bound', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[6,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[6,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[6,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[6,] + Exp1_params[5,]), col='maroon1', xaxs='i', yaxs='i', ylim=c(0, .004), xlab=expression(italic(b)), ylab='', yaxt='n',
     main='Decision\nthreshold', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[6,] + Exp2_params[5,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[6,] + Exp1_params[5,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.5, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[6,] + Exp2_params[5,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.5, col='maroon4', cex=0.75)

plot(density(Exp1_params[1,]), col='maroon1', xaxs='i', yaxs='i', xlim=c(0, 300), ylim=c(0, .06), xlab=expression(italic(t)[0]), ylab='', yaxt='n',
     main='Non-decision\ntime', cex.lab=1.8, cex.main=1.5, las=2)
lines(density(Exp2_params[1,]), col='maroon4', lty=2)
mtext(paste0('M = ', format(round(mean(Exp1_params[1,]), 2), nsmall=2)), side=3, line=-1.5, adj=0.55, col='maroon1', cex=0.75)
mtext(paste0('M = ', format(round(mean(Exp2_params[1,]), 2), nsmall=2)), side=3, line=-2.5, adj=0.55, col='maroon4', cex=0.75)
#dev.off()
```



## Parameter recovery (Model 7: softmax(Q) + gaze)

1. Load parameter recovery results and organize into data frame.

```{r}
paramRecoveryResults <- readRDS('../modeling/results/free-theta/model7_paramRecovery.RDS')

# decision threshold = start point upper bound + threshold separation
paramRecoveryResults[['generating']] <- cbind(paramRecoveryResults[['generating']], 
                                              paramRecoveryResults[['generating']][,5] + 
                                                paramRecoveryResults[['generating']][,6])

paramRecoveryResults[['recovered']] <- cbind(paramRecoveryResults[['recovered']], 
                                              paramRecoveryResults[['recovered']][,5] + 
                                                paramRecoveryResults[['recovered']][,6])

param_names <- c('Learning rate', 'Relative encoding', 'Q drift scaling', 'Gaze drift scaling',
                 'Softmax inv. temp.', 'Start point upper bound', 'Decision threshold', 'Non-decision time')

paramRecoveryData <- data.frame(
  param = factor(rep(param_names, each=100), levels=param_names),
  generating = c(paramRecoveryResults[['generating']][,c(2,7,3,4,8,6,9,1)]),
  recovered = c(paramRecoveryResults[['recovered']][,c(2,7,3,4,8,6,9,1)])
)
```


2. Plot relationships between generating and recovered parameters (SI APPENDIX; S5 Fig)

```{r}
p1 <- ggplot(data=paramRecoveryData, aes(x=generating, y=recovered)) +
  facet_wrap(~ param, nrow=2, scales='free') +
  geom_abline(slope = 1, intercept=0, lty=3) +
  geom_smooth(method='lm', col='maroon1') +
  geom_point(pch=21, fill='grey', alpha=0.5) +
  ylab('Recovered') +
  xlab('True') +
  theme_classic(base_size = 11) +
  theme(panel.grid.major.y = element_blank(),
        plot.title = element_text(hjust=0.5, size=11),
        plot.subtitle = element_text(hjust=0.5, size=11)) +
  ggpubr::stat_cor(aes(label = ..r.label..), method='pearson', cor.coef.name = 'r', cex=3,
                   label.x.npc = .7, label.y.npc = .25) +
  ggtitle('Experiment 2', subtitle='softmax(Q) + gaze model')
```


## Parameter recovery (Model 8: softmax(Q + gaze))

1. Load parameter recovery results and organize into data frame.

```{r}
paramRecoveryResults <- readRDS('../modeling/results/free-theta/model8_paramRecovery.RDS')

# decision threshold = start point upper bound + threshold separation
paramRecoveryResults[['generating']] <- cbind(paramRecoveryResults[['generating']], 
                                              paramRecoveryResults[['generating']][,5] + 
                                                paramRecoveryResults[['generating']][,6])

paramRecoveryResults[['recovered']] <- cbind(paramRecoveryResults[['recovered']], 
                                              paramRecoveryResults[['recovered']][,5] + 
                                                paramRecoveryResults[['recovered']][,6])

param_names <- c('Learning rate', 'Relative encoding', 'Q drift scaling', 'Gaze drift scaling',
                 'Softmax inv. temp.', 'Start point upper bound', 'Decision threshold', 'Non-decision time')

paramRecoveryData <- data.frame(
  param = factor(rep(param_names, each=100), levels=param_names),
  generating = c(paramRecoveryResults[['generating']][,c(2,7,3,4,8,6,9,1)]),
  recovered = c(paramRecoveryResults[['recovered']][,c(2,7,3,4,8,6,9,1)])
)
```


2. Plot relationships between generating and recovered parameters (SI APPENDIX; S5 Fig)

```{r}
p2 <- ggplot(data=paramRecoveryData, aes(x=generating, y=recovered)) +
  facet_wrap(~ param, nrow=2, scales='free') +
  geom_abline(slope = 1, intercept=0, lty=3) +
  geom_smooth(method='lm', col='maroon1') +
  geom_point(pch=21, fill='grey', alpha=0.5) +
  ylab('Recovered') +
  xlab('True') +
  theme_classic(base_size = 11) +
  theme(panel.grid.major.y = element_blank(),
        plot.title = element_text(hjust=0.5, size=11),
        plot.subtitle = element_text(hjust=0.5, size=11)) +
  ggpubr::stat_cor(aes(label = ..r.label..), method='pearson', cor.coef.name = 'r', cex=3,
                   label.x.npc = .7, label.y.npc = .1) +
  ggtitle('Experiment 2', subtitle='softmax(Q + gaze) model')
```


```{r}
# png('Exp2_param_recovery_R1.png', width=7, height=8, units='in', res=300)
# par(mar=c(4.5,1,1.5,1))
# grid.arrange(p1, p2, nrow=2, ncol=1)
# dev.off()
```
